---
title: "Clinseq Cite-seq analysis"
output: html_notebook
---


NB: Code was run on the vale server https://vale.scilifelab.se/ using R v3.6.0 and Seurat v3.2.2 to stick to the original analysis.

Aim : Calculate factor scores based on transcriptomics and compare to AML cell types

Analysis
1. Exploratory analysis : UMAP patients, clusters and normal cell types
2. Calculate cell type score per cell
3. Assign AML cell types based on cell markers
4. Calculate factors scores based on genes expression for each cell
5. Clustering of cells based on factors and cell types
6. Extract ADT values for downstream analysis

### clean environment
```{r}
rm(list=ls())
```

### set directory (replace by your path)
```{r setup}
    knitr::opts_knit$set(root.dir = normalizePath("/home/flomik/Code/MOFA_AML")) 
```

```{r}
library(Seurat)
library(ggpubr)
library(ComplexHeatmap)
library(stringr)
library(xlsx)
library(RColorBrewer)
library(readxl)
library(data.table)
```

## Load CITE-seq (replace to your data folders)
```{r}
data <- readRDS("~/Data/data_francesco/Wells_combined_sctransformed.rds")
```

## Load material
```{r}
path_factors <- "data/AML_MOFAplus_top_factors_v04_f600_2.xlsx"
geneset <- read.csv("data/Genes_sets/Cell_types/Dick/AMLCellType_Genesets.gmt.txt", header = FALSE) ## Dick et al cell markers
vangalen <- read.xlsx("data/Genes_sets/Cell_types/Van_Galen/mmc3.xlsx", 1) ## Van Galen et al cell markers
```

## Load markers from Dick et al paper
```{r}
gs <- list()
for (i in 1:nrow(geneset)) {
  line <- geneset[i,1]
  path <- gsub("\t.*", "", line)
  genes <- gsub(path, "", line)
  genes <- gsub("\t\t", "", genes)
  genes <- str_split(genes, "\t")
  gs[[path]] <- genes[[1]]
  print(path)
}
```

## add Van Galen controls
```{r}
normal_HSC <- as.vector(vangalen$Normal.derived..combined)
normal_GMP <- as.vector(vangalen$NA..1)
normal_myeloid <- as.vector(vangalen$NA..2)
```

```{r}
gs[["HSC_Prog"]] <- normal_HSC[-c(1, 52:54)]
gs[["GMP"]] <- normal_GMP[-c(1, 52:54)]
gs[["Myeloid"]] <- normal_myeloid[-c(1, 52:54)]

print(length(gs))
```

## Calculate score for each cell type per cell (NB cutoff LFC = 3)
```{r}
for (i in 1:length(gs)) {
  features <- list(gs[[i]])
  name_cell <- names(gs)[i]
  print(length(features[[1]][features[[1]] %in% rownames(data@assays$RNA)]))
  data <- AddModuleScore(
    object = data,
    features = features,
    name = name_cell,
    search = TRUE)
}
```

```{r}
print(colnames(data@meta.data)[51:70])
```

```{r}
colnames(data@meta.data)[51:70] <- gsub("1$", "", colnames(data@meta.data)[51:70])
```

# Plot scores of cell types
```{r}
for(i in 51:70){
  print(colnames(data@meta.data)[i])
  
  FeaturePlot(data,
            features = colnames(data@meta.data)[i], label = FALSE, repel = TRUE) +
            scale_colour_gradientn(colours = rev(brewer.pal(n = 11, name = "RdBu")))+ ggplot2::theme(legend.position = "none")
  
  ggsave(paste0("results/figures/score_plot_", colnames(data@meta.data)[i],".pdf"), width = 4, height = 4)
  ggsave(paste0("results/figures/score_plot_", colnames(data@meta.data)[i],".png"), width = 4, height = 4)
}

```

```{r}
write.csv(data@meta.data, "results/scores_Johon_Dick_HC_cells.csv")
```

## remove control cells HL60
```{r}
keep <- rownames(data@meta.data)[!grepl("HL60", data@meta.data$PatientLevel)]
data <- subset(data, cells = keep)
```

```{r}
# Projecting singlet identities on TSNE visualization
DimPlot(data, reduction = "umap", group.by = "Well")+ ggplot2::theme(legend.position = "none")
```

```{r}
col_cells <- c(RColorBrewer::brewer.pal(6,'Set1'), "#d2d5d9")

DimPlot(data, reduction = "umap", group.by = "PatientLevel", cols = col_cells) + ggplot2::theme(legend.position = "none")
ggsave(paste0("results/figures/score_plot_Tojo_patients.pdf"), width = 4, height = 4)
ggsave(paste0("results/figures/score_plot_Tojo_patients.png"), width = 4, height = 4)

table(data@meta.data$PatientLevel)
```


```{r}
DimPlot(data, reduction = "umap", group.by = "predicted.id") + ggplot2::theme(legend.position = "none")
ggsave(paste0("results/figures/score_plot_Tojo_cell_types.png"), width = 4, height = 4)
```

```{r}
DimPlot(data, reduction = "umap", group.by = "seurat_clusters") + ggplot2::theme(legend.position = "none")
ggsave(paste0("results/figures/score_plot_Tojo_clusters.pdf"), width = 4, height = 4)
ggsave(paste0("results/figures/score_plot_Tojo_clusters.png"), width = 4, height = 4)
```

## Assign cells --> extract top 3 scores / only LFC > 3 / score 1 > 0.5 compared to score 2
```{r}
cell_scores <- data@meta.data[,51:70]
cell_scores$score_1 <- NA
cell_scores$score_2 <- NA
cell_scores$score_3 <- NA

cell_scores$cell_1 <- NA
cell_scores$cell_2 <- NA
cell_scores$cell_3 <- NA


topMaxUsingFullSort <- function(x, N) {
  sort(x, decreasing = TRUE)[1:min(N, length(x))]
}


for (i in 1:nrow(cell_scores)) {
  #print(i)
  top_3 <- topMaxUsingFullSort(cell_scores[i, 1:20], 3)
  cell_scores[i, 21] <- top_3[[1]]
  cell_scores[i, 22] <- top_3[[2]]
  cell_scores[i, 23] <- top_3[[3]]
  
  cell_scores[i, 24] <- colnames(top_3)[1]
  cell_scores[i, 25] <- colnames(top_3)[2]
  cell_scores[i, 26] <- colnames(top_3)[3]
}

write.csv(cell_scores, "results/scores_cells_Dick_top_3_assigned_cells.csv")
```

```{r}
cell_scores <- read.csv("results/scores_cells_Dick_top_3_assigned_cells.csv", row.names = 1)
cell_scores <- cell_scores[,c(21:26)]
write.csv(cell_scores, "results/scores_cells_Dick_top_3_assigned_cells_scores_only.csv")
```


## assign cells --> min 3 time more than the random
```{r}
cell_scores <- read.csv("results/scores_cells_Dick_top_3_assigned_cells_scores_only.csv", row.names = 1)

cell_scores$match <- NA
cell_scores$match[cell_scores$score_1 > 2] <- "ok"
table(cell_scores$match)

cell_scores$regex_1 <- cell_scores$cell_1
cell_scores$regex_1 <- gsub("\\..*", "", cell_scores$regex_1)

cell_scores$regex_2 <- cell_scores$cell_2
cell_scores$regex_2 <- gsub("\\..*", "", cell_scores$regex_2)

cell_scores$match[cell_scores$regex_1 == cell_scores$regex_2 & cell_scores$score_1 > 3] <- "db"

cell_scores$score_2[cell_scores$match == "db" & !is.na(cell_scores$match)] <- cell_scores$score_3[cell_scores$match == "db" & !is.na(cell_scores$match)]

cell_scores$diff <- cell_scores$score_1 - cell_scores$score_2

cell_scores$match[cell_scores$diff < 0.5] <- "mix"
cell_scores$match[cell_scores$diff > 0.5 &  cell_scores$score_1 > 3] <- "ok"
table(cell_scores$match)

cell_scores$cell_1[cell_scores$match != "ok"] <- NA

cell_scores$cell_1[cell_scores$score_1 < 2] <- NA

cell_scores$cell_1 <- gsub("\\.Top.*", "", cell_scores$cell_1)

table(cell_scores$match)

table(cell_scores$cell_1)
```

```{r}
data@meta.data$cell_1 <- cell_scores$cell_1
```

```{r}
data@meta.data$cell_1[data@meta.data$cell_1 == "Myeloid" & !is.na(data@meta.data$cell_1) ] <- paste0("Ctrl.", data@meta.data$cell_1[data@meta.data$cell_1 == "Myeloid" & !is.na(data@meta.data$cell_1) ] )

data@meta.data$cell_1[data@meta.data$cell_1 == "GMP" & !is.na(data@meta.data$cell_1) ] <- paste0("Ctrl.", data@meta.data$cell_1[data@meta.data$cell_1 == "GMP" & !is.na(data@meta.data$cell_1) ] )

data@meta.data$cell_1[data@meta.data$cell_1 == "HSC_Prog" & !is.na(data@meta.data$cell_1) ] <- paste0("Ctrl.", data@meta.data$cell_1[data@meta.data$cell_1 == "HSC_Prog" & !is.na(data@meta.data$cell_1) ] )
```

```{r}

col_cells <- c(RColorBrewer::brewer.pal(12,'Paired'), "#d2d5d9")
```

```{r}
data@meta.data$cell_1[is.na(data@meta.data$cell_1)] <- "Unassigned"
```

```{r}
DimPlot(data, reduction = "umap", group.by = "cell_1", cols = c(alpha(col_cells[-13], 1), alpha(col_cells[13], 0.1)))+ ggplot2::theme(legend.position = "none")
ggsave(paste0("results/figures/score_plot_new_cells_V4.pdf"), width = 4, height = 4)
ggsave(paste0("results/figures/score_plot_new_cells_V4.png"), width = 4, height = 4)
```

## calculate factors for each cell
```{r}
#' Perform survival analysis for each MOFA factor
#'
#' @param file variable containing file
#' @param name name of the file
#' @param destination destination folder for file
scores_factor <- function(fname, txn, genes){
  myFactor1=RNAfac[RNAfac$factor==fname,]
  p1=c(1:nrow(myFactor1)) # for coefficients
  
  if(genes == "genes"){
    myFactor1$feature <- myFactor1$label
  }
  
  pick=txn$stable_id %in% myFactor1$feature[p1]
  #print(pick)
  mygex1=txn[pick,drop=FALSE,]
  rownames(mygex1) <- mygex1$stable_id
  mygex1=data.frame(t(mygex1[,colnames(mygex1) != "stable_id"]))
  x=as.matrix(mygex1)
  #print(dim(x))
  
  beta=myFactor1$value[match(colnames(x),myFactor1$feature)]
  beta[is.na(beta)] <- 0
  yval=t(t(x) * beta)
  yval=rowSums(yval)
  
  weight <- data.frame(yval)
  colnames(weight) <- fname
  weight$patient_id <- rownames(weight)
  
  yval <- data.frame(patient_id = names(yval), score = yval)
  yval$group <- as.integer(yval$score > median(yval$score))+1
  return(yval)
}
```

```{r}
gex <- as.data.frame(data@assays$RNA@data)
gex$stable_id <- rownames(gex)
#gex$stable_id <- gsub("\\..*", "", gex$stable_id)
#gex <- aggregate(. ~ stable_id, data = gex, sum)
#rownames(gex) <- gex$stable_id
```

### import factors data
```{r}
RNA_factors=NULL
for (i in 1:11){
  factorName=paste0("Factor",i,"_RNAseq")
  cat(factorName," ")
  dat=read_excel(path_factors,sheet=factorName)
  dat=data.table(dat)
  print(nrow(dat))
  RNA_factors[[i]]=dat
}

RNAfac=do.call(rbind,RNA_factors)
RNAfac$feature=gsub(".g","",RNAfac$feature)

## set thresholds for each factor
factorNames=unique(RNAfac$factor)
topGeneThres=c(9,10,2,6,4,1,2,3,3,6,3)
names(topGeneThres)=factorNames
```

```{r}
table_scores <- data.frame(patient_id = colnames(gex))

for (factor in factorNames){
  print(factor)
  scores_F <- scores_factor(factor, gex, "genes")
  table_scores <- merge(table_scores, scores_F, by = "patient_id")
}

colnames(table_scores)[-1] <- paste0("Factor_", rep(1:11,each = 2), "_",colnames(table_scores)[-1])
colnames(table_scores) <- gsub("\\..*", "", colnames(table_scores))
colnames(table_scores)[1] <- "Cell"
```

```{r}
data@meta.data <- cbind(data@meta.data, table_scores)
```

```{r}
colnames(data@meta.data)[seq(73,93, 2)]
```

```{r}
write.csv(data@meta.data, "results/scores_cells_and_factors_CITE_seq.csv")
```

# Plot scores
```{r}
for(i in seq(73,93, 2)){
  print(colnames(data@meta.data)[i])
  FeaturePlot(data,
            features = colnames(data@meta.data)[i], label = FALSE, repel = TRUE) +
            scale_colour_gradientn(colours = rev(brewer.pal(n = 11, name = "RdBu")))+ ggplot2::theme(legend.position = "none")
  
  ggsave(paste0("results/figures/score_plot_", colnames(data@meta.data)[i],"_scores_MOFA.pdf"), width = 4, height = 4)
  ggsave(paste0("results/figures/score_plot_", colnames(data@meta.data)[i],"_scores_MOFA.png"), width = 4, height = 4)
}

```


```{r}
cells <- data@meta.data[!is.na(data@meta.data$cell_1), ]
cells$cell_1 <- as.vector(cells$cell_1)
x <- table(cells$cell_1) > 50
x <- names(x)[x == TRUE]
```

```{r}
cells <- data@meta.data[data@meta.data$cell_1 %in% x, ]
cells$cell_1 <- as.vector(cells$cell_1)
table(cells$cell_1)
```

```{r}
write.csv(data@meta.data, "results/scores_cells_and_factors_CITE_seq_clean.csv")
```


```{r}
my_comparisons <- list(c("LSPC.Cycle", "LSPC.Quiescent"), c("cDC.like", "GMP.like."), c("cDC.like", "LSPC.Quiescent"), c("GMP.like", "LSPC.Cycle"), c("GMP.like", "LSPC.Quiescent"), c("cDC.like", "LSPC.Cycle"))

for (i in seq(73,93, 2)) {
  print(colnames(cells)[i])
  gg <- ggplot(cells, aes(y=cells[,i], x=cell_1, fill = cell_1)) + 
  geom_boxplot()+theme(legend.position = "none")+ 
  stat_compare_means(comparisons = my_comparisons)+theme_classic()+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
ggsave(paste0("results/figures/boxplots_cell_types_", colnames(cells)[i],".pdf"), height = 6, width = 8)
ggsave(paste0("results/figures/boxplots_cell_types_", colnames(cells)[i],".png"), height = 6, width = 8)
gg
}
```

```{r}
cells <- cells[cells$cell_1 != "Unassigned",]
colnames(cells)[seq(73,93, 2)] <- gsub("Factor_", "DDHM", colnames(cells)[seq(73,93, 2)])
colnames(cells)[seq(73,93, 2)] <- gsub("_score", "", colnames(cells)[seq(73,93, 2)])
```

```{r}
cells[,seq(73,93, 2)] <- scale(cells[,seq(73,93, 2)] )
```

```{r}
column_ha <- HeatmapAnnotation(cell_type = cells$cell_1)

pdf("results/figures/heatmap_factors_scaled_cell_types_clustering.pdf")
Heatmap(t(cells[,seq(73,93, 2)]), show_row_names = TRUE, show_column_names = FALSE, row_title_rot = 0, top_annotation = column_ha)
dev.off()
```

```{r}
pdf("results/figures/heatmap_factors_scaled_cell_types.pdf")
Heatmap(cells[,seq(73,93, 2)], row_split = cells$cell_1, show_row_names = FALSE, row_title_rot = 0)
dev.off()
```

```{r}
data_adt <- data@assays$ADT@scale.data
```

```{r}
write.csv(data_adt, "results/data_ADT_cell_markers_norm.csv")
```


```{r}
#saveRDS(data, file = "data/Wells_combined_sctransformed_with_Dick_cell_types.rds")
```


```{r}
sessionInfo()
```

