---
title: "AML preprocess for MOFA+ v04"
output:
  html_document:
    fig_width: 9
    fig_height: 7
    theme: cerulean
    toc: yes
    number_sections: true
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

Require all the libraries first
```{r message=F}
require(MultiAssayExperiment)
require(SummarizedExperiment)
require(minfi)
require(DESeq2)
require(MOFA2)
require(RColorBrewer)
require(tidyverse)
```

# Load the data

```{r}
load("/AML/data/MultiAssayExp/AML_MAexp_v05.RData")
outDir <-  "/AML/out/MOFA_v04"
dir.create(outDir)
```

# Preprocess

## RNAseq

Remore features with low counts, VST transform, exclude genes on Y chromosome and select the top 5000 variable features.

```{r}
idx <- rowSums(counts(MAexp[["RNAseq"]], normalized=TRUE) >= 10 ) >= ncol(MAexp[["RNAseq"]])/2
dds.filter <- MAexp[["RNAseq"]][idx,]

dds.filter <- estimateSizeFactors(dds.filter)
vst.filter <- varianceStabilizingTransformation(dds.filter)

geneplotter::multidensity(assay(vst.filter), xlab="VST", lwd=2, legend = NULL, main="VST-transform")

vst.filter <- vst.filter[!rownames(vst.filter) %in% na.omit(rownames(vst.filter)[mcols(vst.filter)$chromosome_name=="Y"])]

feat.var <- rowVars(assay(vst.filter))
names(feat.var) <- rownames(vst.filter) 

idx <- rownames(vst.filter) %in% names(sort(feat.var, decreasing = T))[1:5000]
vst.filter <- vst.filter[idx,]

geneplotter::multidensity(assay(vst.filter), xlab="VST", lwd=2, legend = NULL, main="VST-transform")
```

## Proteomics

Consider features detected in all samples, exclude genes on Y chromosome and select the top 5000 variable features.

```{r}
idx <- rowSums(is.na(assay(MAexp[["Proteomics"]])))==0
prot.se.filter <- MAexp[["Proteomics"]][idx,]

geneplotter::multidensity(assay(prot.se.filter), xlab="log2R", lwd=2, legend = NULL)

prot.se.filter <- prot.se.filter[!rownames(prot.se.filter) %in% na.omit(rownames(prot.se.filter)[mcols(prot.se.filter)$chromosome_name=="Y"])]

feat.var <- rowVars(assay(prot.se.filter))
names(feat.var) <- rownames(prot.se.filter) 

idx <- rownames(prot.se.filter) %in% names(sort(feat.var, decreasing = T))[1:5000]
prot.se.filter <- prot.se.filter[idx,]

geneplotter::multidensity(assay(prot.se.filter), xlab="log2R", lwd=2, legend = NULL)
```

## DSRT

```{r}
geneplotter::multidensity(assay(MAexp[["DSRT"]]), xlab="sDSS", lwd=2, legend = NULL)
```

## Mutations

Remove mutations detected in less than 3 samples

```{r}
mut.se.filter <- MAexp[["Mutations"]][rowSums(assays(MAexp[["Mutations"]])$mut.pres)>=3,]
tmp <- assays(mut.se.filter)$mut.pres

# tmp.dend <- hclust(dist(t(tmp), method = "binary"))
# tmp.dend <- reorder(as.dendrogram(tmp.dend), wts = colSums(tmp), agglo.FUN = max)

tmp[tmp==1] <- "Mut"
tmp[tmp==0] <- ""

col=c(Mut="blue")

ComplexHeatmap::oncoPrint(tmp, get_type = function(x) {x},
                          
                          alter_fun = list(
                            Mut = function(x, y, w, h) grid::grid.rect(x, y, w*0.9, h*0.9, gp = grid::gpar(fill = col["Mut"], col = NA))
                          ),
                          col=col, show_pct = F,
                          row_order = rownames(tmp)[order(rowSums(assays(mut.se.filter)$mut.pres), decreasing = T)])

```

## Cytogenetic alterations

Remove alterations detected in less than 3 samples

```{r}
cyt.se.filter <- MAexp[["Cytogenetics"]][rowSums(assay(MAexp[["Cytogenetics"]]))>=3,]
tmp <- assay(cyt.se.filter)

tmp[tmp==1] <- "Alt"
tmp[tmp==0] <- ""

col=c(Alt="blue")

ComplexHeatmap::oncoPrint(tmp, get_type = function(x) {x},
                          
                          alter_fun = list(
                            Alt = function(x, y, w, h) grid::grid.rect(x, y, w*0.9, h*0.9, gp = grid::gpar(fill = col["Alt"], col = NA))
                          ),
                          col=col, show_pct = F,
                          row_order = rownames(tmp)[order(rowSums(assay(cyt.se.filter)), decreasing = T)])

```

## DNAm

Exclude probes on X and Y chromosomes and select the top 5000 variable features. Transform to M values.

```{r}
data.Rset.filt <- MAexp[["DNAm"]][as.logical(!seqnames(MAexp[["DNAm"]])=="chrX" & !seqnames(MAexp[["DNAm"]])=="chrY"),]

feat.var <- rowVars(assays(data.Rset.filt)$Beta)
names(feat.var) <- rownames(data.Rset.filt) 

idx <- rownames(data.Rset.filt) %in% names(sort(feat.var, decreasing = T))[1:5000]
data.Rset.filt <- data.Rset.filt[idx,]

assays(data.Rset.filt)$M <- getM(data.Rset.filt)
 
geneplotter::multidensity(assays(data.Rset.filt)$Beta, xlab="Beta", lwd=2, legend = NULL)
geneplotter::multidensity(assays(data.Rset.filt)$M, xlab="M", lwd=2, legend = NULL)

```


# Create MOFA object

A data frame with columns "sample","feature","view","group","value" needs to be created and saved as text file. The mutations and the cytogenetics alterations will be on the same view ("Mutations")

```{r}
assay(vst.filter) %>%
  as.data.frame() %>% 
  rownames_to_column("feature.ID") %>%
  mutate(feature = paste0(sapply(strsplit(feature.ID, ".", fixed = T), "[",1), ".g")) %>% 
  gather(key="sample", value = "value", -c(feature, feature.ID)) %>% 
  mutate(view="RNAseq",
         group="AML") %>% 
  select(sample, feature, view, group, value) -> g.df

assay(prot.se.filter) %>% 
  as.data.frame() %>% 
  rownames_to_column("feature.ID") %>%
  mutate(feature=paste0(rowData(prot.se.filter)$ensembl_gene_id, ".p")) %>%
  filter(feature!="NA.p") %>% 
  gather(key="sample", value = "value", -c(feature, feature.ID)) %>% 
  mutate(view="Proteomics",
         group="AML") %>% 
  select(sample, feature, view, group, value) -> p.df

assay(assay(MAexp[["DSRT"]])) %>% 
  as.data.frame() %>% 
  rownames_to_column("feature") %>%
  gather(key="sample", value = "value", -feature) %>% 
  mutate(view="DSRT",
         group="AML") %>% 
  select(sample, feature, view, group, value) -> d.df


assay(assays(mut.se.filter)$mut.pres) %>% 
  as.data.frame() %>% 
  rownames_to_column("feature") %>%
  gather(key="sample", value = "value", -feature) %>% 
  mutate(view="Mutations",
         group="AML") %>% 
  select(sample, feature, view, group, value) -> m.df

assay(assay(cyt.se.filter)) %>% 
  as.data.frame() %>% 
  rownames_to_column("feature") %>%
  gather(key="sample", value = "value", -feature) %>% 
  mutate(view="Mutations",
         group="AML") %>% 
  select(sample, feature, view, group, value) -> cyt.df


assay(assays(data.Rset.filt)$M) %>% 
  as.data.frame() %>% 
  rownames_to_column("feature") %>%
  gather(key="sample", value = "value", -feature) %>% 
  mutate(view="DNAm",
         group="AML") %>% 
  select(sample, feature, view, group, value) -> DNAm.df
```

Create annotation data.frame.

```{r}
tmp1 <- data.frame(feature = unique(g.df$feature), stringsAsFactors = F)
tmp1$id <- sapply(strsplit(tmp1$feature, split = ".", fixed = T),"[",1)
tmp1$name <- rowData(dds.filter)$hgnc_symbol[match(tmp1$id, rowData(dds.filter)$ensembl_gene_id)]

tmp2<- data.frame(feature = unique(p.df$feature), stringsAsFactors = F)
tmp2$id <- sapply(strsplit(tmp2$feature, split = ".", fixed = T),"[",1)
tmp2$name <- rowData(prot.se.filter)$hgnc_symbol[match(tmp2$id, rowData(prot.se.filter)$ensembl_gene_id)]

tmp3 <- data.frame(feature = unique(d.df$feature), stringsAsFactors = F)
tmp3$id <- tmp3$feature
tmp3$name <- rowData(MAexp[["DSRT"]])$feature.name[match(tmp3$id, rownames(MAexp[["DSRT"]]))]

tmp4 <- data.frame(feature = unique(m.df$feature), stringsAsFactors = F)
tmp4$id <- tmp4$feature
tmp4$name <- tmp4$feature

tmp5 <- data.frame(feature = unique(cyt.df$feature), stringsAsFactors = F)
tmp5$id <- tmp5$feature
tmp5$name <- rowData(cyt.se.filter)$name[match(tmp5$id, rownames(cyt.se.filter))]

annot.Rset <-getAnnotation(data.Rset.filt)
tmp6 <- data.frame(feature = unique(DNAm.df$feature), stringsAsFactors = F)
tmp6 %>% 
  mutate(id = feature,
         name = annot.Rset$UCSC_RefGene_Name[match(feature, annot.Rset$Name)]) %>%
  mutate(name = sapply(name, function(x) paste(sapply(str_split(x,";"),unique),collapse = ";"))) -> tmp6

annot.df <- rbind(tmp1, tmp2, tmp3, tmp4, tmp5, tmp6)
rm(tmp1, tmp2, tmp3, tmp4, tmp5, tmp6)
```


Create MOFA data

```{r}

mofa.data <- create_mofa(rbind(g.df, p.df, d.df, m.df, cyt.df, DNAm.df))
mofa.data
plot_data_overview(mofa.data)
```

Write data to txt file.

```{r}
write.table(rbind(g.df, p.df, d.df, m.df, cyt.df, DNAm.df),
            file="~/Projects/AML/out/MOFA_v04/MOFA_v04_input.txt", sep="\t", quote=F, row.names = F)
write.table(annot.df,
            file="~/Projects/AML/out/MOFA_v04/MOFA_v04_annot_features.txt", sep="\t", quote=F, row.names = F)
```

