---
title: "Downstream analysis MOFA"
output:
  pdf_document: default
  html_notebook: default
---

NB: Code was run on the vale server https://vale.scilifelab.se/ using R v3.6.0 and MOFA2 v0.99.7 to stick to the original analysis.

Aim : Quality check model and enrichment of pathways and cell types

Analysis:
1. Visualization and model quality
2. Correlation factors with clinical data
3. Pathway analysis for each factor
4. Cell type enrichment for each factor

### clean environment
```{r}
rm(list=ls())
```

### set directory (replace by your repository)
```{r setup}
    knitr::opts_knit$set(root.dir = normalizePath("/home/flomik/Code/MOFA_AML/")) 
```

### load libraries
```{r}
library(MultiAssayExperiment)
library(MOFA2)
library(RColorBrewer)
library(GSEABase)
library(tidyverse)
library(minfi)
library(xlsx)
library(org.Hs.eg.db)
library(maditr)
library(psych)
```

# Analysis

## load data (replace by your repository)
```{r}
## MOFA
load("data/input_files/MOFA_input/AML_MAexp_v05_filt.RData")
load("data/input_files/MOFA_object/MOFA_v04_train_list.RData")
ls()

## clinical
clinical_data <- pData

## pathways
ensembl.react <- read.table("data/Pathways/Ensembl2Reactome_All_Levels.txt", header = F, sep="\t", comment.char ="", quote = "")
```

## load annotations
```{r}
annot.df <- read.table("data/MOFA/MOFA_v04_annot_features.txt", sep="\t", header = T, stringsAsFactors = F)
```

## Select MOFA object
```{r}
mofa.sel <- select_model(mofa.list)
```

## Model overview.
```{r}
plot_data_overview(mofa.sel)
pdf("results/figures/MOFA/MOFA_overview_model_Tojo.pdf")
plot_data_overview(mofa.sel)
dev.off()
```

## Correlation factors
Verify that the Factors are largely uncorrelated
```{r}
plot_factor_cor(mofa.sel)
pdf("results/figures/MOFA/correlation_factors_model_Tojo.pdf")
plot_factor_cor(mofa.sel)
dev.off()
```

## variance decomposition by factor
Percentage of variance explained by each factor across each data modality
```{r}
plot_variance_explained(mofa.sel)
pdf("results/figures/MOFA/MOFA_variance_model_Tojo.pdf")
plot_variance_explained(mofa.sel)
dev.off()
```

## Total variance
model is providing a good fit to the data
```{r}
plot_variance_explained(mofa.sel, plot_total = T)[[2]]
pdf("results/figures/MOFA/MOFA_variance_modeL_Tojo_plot_2.pdf")
plot_variance_explained(mofa.sel, plot_total = T)[[2]]
dev.off()
```

## Checking model / percentage variance per factor
```{r}
mofa.sel@cache
```

```{r}
colnames(mofa.sel@samples_metadata)
```

## filter and order clinical data
```{r}
clinical_data <- clinical_data[!duplicated(clinical_data$patient_id),]
names(clinical_data)[1] <- "sample"
clinical_data$group <- "AML"
table(clinical_data$ELN_2022_AO)
clinical_data <- clinical_data[order(samples_names(mofa.sel)$AML, clinical_data$sample),]
```

## add metadata
```{r}
# Sanity check
stopifnot(all(sort(clinical_data$sample)==sort(unlist(samples_names(mofa.sel)))))

# Add sample metadata to the model
samples_metadata(mofa.sel) <- clinical_data
```

```{r}
clinical_data <- clinical_data[,!grepl("date", colnames(clinical_data))]
clinical_data$vitally_frozen_million_cells <- NULL
clinical_data$karyotype <- NULL
clinical_data$karyotype_info <- NULL
clinical_data$aml_subtype <- NULL
clinical_data$group <- NULL
clinical_data$MDS <- NULL
clinical_data$MPD <- NULL
clinical_data$sample_type <- NULL
```

```{r}
clinical_data$sample
```

```{r}
clinical_data$treat_intention <- NULL
```

## relationship clinical data and latent factors
```{r}
factors <- data.frame(sample = rownames(mofa.sel@expectations$Z$AML), mofa.sel@expectations$Z$AML)
factors_clinical <- merge(clinical_data, factors, by = "sample")
total_regression_table <- as.data.frame(matrix(0,nrow=0,ncol=7))
colnames(total_regression_table) <- c("Clinical_parameter","Factor","Estimate","StdError","tStat","pvalue", "padjust")

for (i in 2:26) {
  for(j in 27:37){
    plm <- as.data.frame(matrix(0,nrow=1,ncol=6))
    colnames(plm) <- c("Clinical_parameter","Factor","Estimate","StdError","tStat","pvalue")
    #print(names(factors_clinical)[i])
    plm[1, 1] <- names(factors_clinical)[j]
    plm[1, 2] <- names(factors_clinical)[i]
    plm[1,3:6] <- coef(summary(lm(factors_clinical[,j] ~ factors_clinical[,i], data = factors_clinical)))[2,]
    plm$padjust <- p.adjust(plm$pvalue,method="BH")
    total_regression_table <- rbind(total_regression_table, plm)
  }
}

sigplm <- total_regression_table
#sigplm <- total_regression_table[total_regression_table$padjust<0.05,] #look only at those with adjusted p.value < 0.05
write.csv(sigplm, "results/MOFA/association_clinical_MOFA_latent_factors.csv")
```

```{r}
sign <- total_regression_table[total_regression_table$padjust<0.05,]
write.csv(sign, "results/MOFA/association_clinical_MOFA_latent_factors_filt.csv")
```

```{r}
estimates <- dcast(sigplm, Clinical_parameter ~ Factor, value.var = "Estimate")
row_names_table <- estimates$Clinical_parameter
estimates$Clinical_parameter <- NULL
estimates <- data.frame(estimates)
#estimates[abs(estimates) < 0.5]  <- NA
```

```{r}
p.values <- dcast(sigplm, Clinical_parameter ~ Factor, value.var = "padjust")

row_names_table <- p.values$Clinical_parameter
p.values$Clinical_parameter <- NULL
p.values <- data.frame(p.values)
p.values[p.values > 0.05] <- NA
#p.values[abs(estimates) < 0.2]  <- NA
p.values <- -log10(p.values)
p.values <- t(p.values)
colnames(p.values) <- row_names_table

rownames(p.values)

p.values <- p.values[,c(1, 4, 5, 6, 7, 8, 9, 10, 11, 2, 3)]
p.values <- p.values[c(8, 9, 10, 11, 12, 13, 14, 15, 16, 1, 2, 17, 3, 18, 4, 5, 6, 19, 20, 21, 22, 23, 24, 7, 25),]

#p.values_2 <- p.values
#p.values_2[8:9,1] <- "no_Tojo"
#p.values_2[!is.na(p.values_2[,1]),2] <- "no_Tojo"
#p.values_2[1:9,3] <- "no_Tojo"
#p.values_2[1:9,3] <- "no_Tojo"

col <- colorRampPalette(c("grey97","red"))(n = 100)
pheatmap::pheatmap(p.values, color = col, show_rownames = T, labels_row = str_trunc(rownames(p.values), width = 40),
                   cluster_rows=FALSE, cluster_cols=FALSE,border_color = "#555555",fontsize =14,fontsize_row=14,fontsize_col=14)

pdf("results/figures/MOFA/heatmap_clinical_association_fators.pdf")
pheatmap::pheatmap(p.values, color = col, show_rownames = T, labels_row = str_trunc(rownames(p.values), width = 40),
                   cluster_rows=FALSE, cluster_cols=FALSE,border_color = "#555555",fontsize =14,fontsize_row=14,fontsize_col=14)
dev.off()
```


## Correlation with clinical data
```{r}
correlate_factors_with_covariates(mofa.sel, 
  covariates = c("sex", "age", "aml_etiology", "blast_bm_percentage", "blast_blood_abs", "genetic_risk", "complete_remission", "status", "ELN_2022_AO", "cytogenetic_risk", "who_perfstatus", "ELN2017", "time_days"), 
  plot="log_pval"
)
```
