---
title: "survival analysis MOFA using MOFA scores based on transcriptomics"
output: html_notebook
---

Aim : calculate scores based on transcriptomics and survival analysis between low and high score value 

Analysis
1. Calculate factors based on linear combination of top 600 genes for each factor in swedish cohort
2. Survival analysis using Factor 8 (high/low) in swedish cohort
3. Survival analysis using Factor 8 (high/low) calculated from transcriptomics in swedish cohort

4. Calculate factors based on linear combination of top 600 genes for each factor in BeatAML cohort
5. Survival analysis using Factor 8 (high/low) calculated from transcriptomics in BeatAML cohort (chemotherapy)
6. Survival analysis using Factor 4 (high/low) calculated from transcriptomics in BeatAML cohort (sorafenib)
7. Survival analysis using Factor 2 (high/low) calculated from transcriptomics in BeatAML cohort (ven+aza)

### clean environment
```{r}
rm(list=ls())
```

```{r}
path <- getwd()
new_path <- gsub("MOFA_AML/.*","MOFA_AML/",path)
```

### set directory (replace by your repository)
```{r setup}
    knitr::opts_knit$set(root.dir = normalizePath(new_path)) 
```

```{r}
library(MOFA2)
library(data.table)
library(readxl)
library(ggplot2)
library(survival)
library(survminer)
library("smplot2")
library(MultiAssayExperiment)
library(gridExtra)
library(xlsx)
```

## load data
```{r}
load("data/input_files/MOFA_input/AML_MAexp_v05.RData")


load("data/input_files/MOFA_object/MOFA_v04_train_list.RData")
```

## load annotations
```{r}
annot.df <- read.table("data/MOFA/MOFA_v04_annot_features.txt", sep="\t", header = T, stringsAsFactors = F)
```

## Select MOFA object
```{r}
mofa.sel <- select_model(mofa.list)
```


## extract top features
```{r}
path_factors <- "data/MOFA/AML_MOFAplus_top_factors_v04_f600 (1).xlsx"
RNA_factors=NULL
for (i in 1:11){
  factorName=paste0("Factor",i,"_RNAseq")
  cat(factorName," ")
  dat=read_excel(path_factors,sheet=factorName)
  dat=data.table(dat)
  print(nrow(dat))
  RNA_factors[[i]]=dat
}
RNAfac=do.call(rbind,RNA_factors)
RNAfac$feature=gsub(".g","",RNAfac$feature)
## set thresholds for each factor
factorNames=unique(RNAfac$factor)
topGeneThres=c(9,10,2,6,4,1,2,3,3,6,3)
names(topGeneThres)=factorNames
```

## Extract transcriptomics
```{r}
gex <- data.frame(mofa.sel@data[[5]]$AML)
gex <-gex[!is.na(gex),]
rownames(gex) <- gsub(".g", "", rownames(gex))
gex$stable_id <- rownames(gex)
txn <- gex[!is.na(gex$ALBB13002),]
txn$ALG200931 <- NULL
condition <- data.frame(patient_id = colnames(txn))
```

## Calculate factor scores transcriptomics
```{r}
for (i in 1:11) {
  fname <- paste0("Factor", i)
  myFactor1 <- RNAfac[RNAfac$factor==fname,]
  pick <- txn$stable_id %in% myFactor1$feature
  x=txn[pick,drop=FALSE,]
  x=t(x[,colnames(x) != "stable_id"])
  beta <- myFactor1$value[match(colnames(x), myFactor1$feature)]
  yval=t(t(x) * beta)
  yval=rowSums(yval)
  weight <- data.frame(yval)
  colnames(weight) <- fname
  weight$patient_id <- rownames(weight)
  yval <- data.frame(patient_id = names(yval), score = yval)
  yval$group <- as.integer(yval$score > median(yval$score))+1
  colnames(yval)[2:3] <- paste0(fname, "_", colnames(yval)[2:3])
  condition <- merge(condition, yval, by = "patient_id")
}
```

```{r}
write.csv(condition, "processing/MOFA_transcriptomics_scores.csv")
```

## extract DDHM values
```{r}
factors <- read_xlsx("data/MOFA/DDHMpatValue.xlsx")
patients <- factors$patient_id
rownames(factors) <- factors$patient_id
factors$sample <- NULL
factors$patient_id <- NULL
factors <- data.frame(lapply(factors, as.numeric))
factors$patient_id <- patients
txn <- condition[,colnames(condition) %in% c("patient_id", colnames(condition)[grepl("score", colnames(condition))])]
data <- merge(factors, txn, by = "patient_id")
```

```{r}
data$Factor4_score <- -1*data$Factor4_score
data$Factor8_score <- -1*data$Factor8_score
```

```{r}
myplots <- vector('list', 11)

for (i in 2:12) {
  myplots[[i-1]] <- local({
        i <- i
        a <- ggplot(data = data, mapping = aes(x = data[, i], y =data[, i+11])) +
          geom_point(size = 1) + xlab("") + ylab("")+theme_classic()+
          sm_statCorr( color = "#0f993d", corr_method = "pearson", linetype = "dashed")+
          ggtitle(paste0("DDHM", i-1))
        print(a)
    })
}
```
```{r}
colnames(data)[5]
```

```{r}
colnames(data)[5+11]
```
```{r}
a <- ggplot(data = data, mapping = aes(x = data[, 3], y =data[, 3+11])) +
          geom_point(size = 1) + xlab("DDHM2 (MOFA)") + ylab("DDHM2 (Transcriptomics score)")+
          sm_statCorr(corr_method = "pearson", size = 0.1) +
  geom_smooth(method=lm, level=0.95, colour = "#4076A9", fill = "#4076A9", alpha=0.25)+
  theme_light()+ 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                                      text=element_text(family="Helvetica", size=12))
a
ggsave("results/figures/Associations/scatter_plots_MOFA_vs_transcriptomics_Factor_2_v2.pdf", width = 4, height = 4, units = "in")
```
,
```{r}
a <- ggplot(data = data, mapping = aes(x = data[, 5], y =data[, 5+11])) +
          geom_point(size = 1) + xlab("DDHM4 (MOFA)") + ylab("DDHM4 (Transcriptomics score inverted sign)")+
          sm_statCorr(corr_method = "pearson", size = 0.1) +
  geom_smooth(method=lm, level=0.95, colour = "#BED797", fill = "#BED797", alpha=0.25)+
  theme_light()+ 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                                      text=element_text(family="Helvetica", size=12))
a
ggsave("results/figures/Associations/scatter_plots_MOFA_vs_transcriptomics_Factor_4_v2.pdf", width = 4, height = 4, units = "in")
```
,


```{r}
a <- ggplot(data = data, mapping = aes(x = data[, 9], y =data[, 9+11])) +
          geom_point(size = 1) + xlab("DDHM8 (MOFA)") + ylab("DDHM8 (Transcriptomics score inverted sign)")+
          sm_statCorr(corr_method = "pearson", size = 0.1) +
  geom_smooth(method=lm, level=0.95, colour = "#FF7F00", fill = "#FF7F00", alpha=0.25)+
  theme_light()+ 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                                      text=element_text(family="Helvetica", size=12))
a
ggsave("results/figures/Associations/scatter_plots_MOFA_vs_transcriptomics_Factor_8_v2.pdf", width = 4, height = 4, units = "in")
```

```{r}
plots <- marrangeGrob(myplots, nrow = 4, ncol = 3, layout_matrix = matrix(
    seq_len(4 * 3),
    nrow = 4, ncol = 3,
    byrow = TRUE
  ))
ggsave("results/figures/Associations/scatter_plots_MOFA_vs_transcriptomics_v2.pdf", plots, width = 8, height = 8, units = "in")
```

## load clinical data
```{r}
clinical_data <- data.frame(pData)
clinical_data <- clinical_data[, colnames(clinical_data) %in% c("time_days", "status", "patient_id")]
clinical_data$survival <- clinical_data$time_days/365
data <- merge(factors, condition, by = "patient_id")
data <- merge(clinical_data, data, by = "patient_id")
```


## separate groups between high Factor 8 and low Factor 8
```{r}
data$Factor8_group <- paste0("G", data$Factor8_group)
data$Factor8_group_2 <- NA
data$status
data$DDHM8_group <- NA
```


```{r}
data$DDHM8_group[data$DDHM8 > median(data$DDHM8)] <- 2
data$DDHM8_group[data$DDHM8 < median(data$DDHM8)] <- 1
```


```{r}
data$Factor8_score <- -1*data$Factor8_score
```

```{r}
data$Factor8_group_2[data$Factor8_score > median(data$Factor8_score)] <- 2
data$Factor8_group_2[data$Factor8_score < median(data$Factor8_score)] <- 1
```

```{r}
data$Factor8_group_2[is.na(data$Factor8_group_2)] <- 2
```

## Survival analysis
```{r}
# run cox regression based on groups
fit.surv <- survfit(Surv(survival, status) ~ DDHM8_group, data = data)

ggsurv <- ggsurvplot(fit.surv,
           data = data,
           risk.table = TRUE,
           conf.int = TRUE,
           size = 1,
           pval = TRUE,
           palette = c("#0000FF", "#FB0207"),
           legend.labs = c("m_lo", "m_hi"),
           ggtheme = theme_light(),
           tables.height = 0.23,
           xlim = c(0, 15),
           pval.coord = c(0,0),
           pval.size = 4.5,
           conf.int.alpha = 0.25)

ggsurv$plot <- ggsurv$plot + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text=element_text(family="Helvetica", size=12)) +xlab("Time (years)")+ 
  theme(strip.text.x = element_blank(),
      strip.background = element_rect(colour="white", fill="white"),
      legend.position=c(.85,.85))+ggtitle("")

ggsurv$table <- ggsurv$table + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                                      text=element_text(family="Helvetica", size=12))

ggsurv 
pdf("results/figures/Survival/survival_curves_DDHM8_median.pdf", width = 4, height = 5)
print(ggsurv, newpage = FALSE)
dev.off()
```

```{r}
# run cox regression based on groups
fit.surv <- survfit(Surv(survival, status) ~ Factor8_group_2, data = data)

ggsurv <- ggsurvplot(fit.surv,
           data = data,
           risk.table = TRUE,
           conf.int = TRUE,
           size = 1,
           pval = TRUE,
           palette = c("#0000FF", "#FB0207"),
           legend.labs = c("t_lo", "t_hi"),
           ggtheme = theme_light(),
           tables.height = 0.23,
           xlim = c(0, 15),
           pval.coord = c(0,0),
           pval.size = 4.5,
           conf.int.alpha = 0.25)

ggsurv$plot <- ggsurv$plot + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text=element_text(family="Helvetica", size=12)) +xlab("Time (years)")+ 
  theme(strip.text.x = element_blank(),
      strip.background = element_rect(colour="white", fill="white"),
      legend.position=c(.85,.85))

ggsurv$table <- ggsurv$table + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                                      text=element_text(family="Helvetica", size=12))+ggtitle("")

ggsurv
pdf("results/figures/Survival/survival_curves_DDHM8_scores_txn_KI_median_inverted_sign.pdf", width = 4, height = 5)
print(ggsurv, newpage = FALSE)
dev.off()
```
## extract sDSS

## Survival DDHM8 BeatAML

### paths files / analysis
```{r}
name_ana <- "BeatAML"
path_txn <- "data/input_files/BeatAML/beataml_waves1to4_norm_exp_dbgap.txt"
path_clinical <- "data/BeatAML/beataml_wv1to4_clinical.xlsx"
path_factors <- "data/MOFA/AML_MOFAplus_top_factors_v04_f600 (1).xlsx"
```

/home/flomik/Data/external_data_sets/BeatAML/beataml_waves1to4_norm_exp_dbgap.txt
## add categories
```{r}
categories_drug <- read.xlsx("data/Drugs/Annotation_curated_DrugBank_ChEMBL_class_revised_2021-10-28.xlsx", 1)
```

### import gene expression
```{r}
gex <- read.delim(path_txn,  row.names = NULL)
gex$display_label <- NULL
rownames(gex) <- gex$stable_id
gex$stable_id <- NULL
```

### import clinical data and preprocessing (rename columns, remove samples, harmonize, make sub-groups) / different for each cohort
```{r}
clinics <- read.xlsx(path_clinical, 1)
colnames(clinics)[colnames(clinics) == "vitalStatus"] <- "status"
colnames(clinics)[colnames(clinics) == "overallSurvival"] <- "follow_up_days"
colnames(clinics)[colnames(clinics) == "dbgap_rnaseq_sample"] <- "patient_id"

table(clinics$status)
## remove patients with unknow status
clinics <- clinics[clinics$status != "Unknown",]
clinics$status=ifelse(clinics$status=="Dead", 1, 0)

## calculate survival
clinics$survival <- clinics$follow_up_days/365
clinics <- clinics[!duplicated(clinics$patient_id),]
clinics <- clinics[!is.na(clinics$patient_id),]
rownames(clinics) <- clinics$patient_id

clinics$ELN2017_char <- clinics$ELN2017
clinics$ELN2017[!clinics$ELN2017_char %in% c("Adverse", "Favorable", "Intermediate")] <- NA

clinics$ELN2017[clinics$ELN2017 == "Favorable" & !is.na(clinics$ELN2017)] <- 1
clinics$ELN2017[clinics$ELN2017 == "Intermediate" & !is.na(clinics$ELN2017)] <- 2
clinics$ELN2017[clinics$ELN2017 == "Adverse" & !is.na(clinics$ELN2017)] <- 3
clinics$ELN2017 <- as.numeric(clinics$ELN2017)
table(clinics$ELN2017)
```

### matching between clinics and gene expression
```{r}
mypatients=intersect(clinics$patient_id, colnames(gex))
length(mypatients)
gex=gex[,c(match(mypatients,colnames(gex)))]
clinics=clinics[match(mypatients,clinics$patient_id),]
sum(clinics$patient_id!=colnames(gex)) #check if they are the same order
```

## Extract transcriptomics
```{r}
txn <- gex
txn$stable_id <- rownames(txn)
condition <- data.frame(patient_id = colnames(txn))
```

## Calculate scores transcriptomics
```{r}
for (i in 1:11) {
  fname <- paste0("Factor", i)
  myFactor1 <- RNAfac[RNAfac$factor==fname,]
  pick <- txn$stable_id %in% myFactor1$feature
  x=txn[pick,drop=FALSE,]
  x=t(x[,colnames(x) != "stable_id"])
  beta <- myFactor1$value[match(colnames(x), myFactor1$feature)]
  yval=t(t(x) * beta)
  yval=rowSums(yval)
  weight <- data.frame(yval)
  colnames(weight) <- fname
  weight$patient_id <- rownames(weight)
  yval <- data.frame(patient_id = names(yval), score = yval)
  yval$group <- as.integer(yval$score > median(yval$score))+1
  colnames(yval)[2:3] <- paste0(fname, "_", colnames(yval)[2:3])
  condition <- merge(condition, yval, by = "patient_id")
}
```

```{r}
write.csv(condition, "processing/scores_BeatAML_MOFA_transcriptomics.csv")
```

```{r}
clinical =clinics[,colnames(clinics) %in% c("patient_id","status", "survival")]
patients_chemo <- clinics$patient_id[grepl("Standard Chemotherapy", clinics$typeInductionTx)]
data <- merge(condition, clinical, by = "patient_id")
data <- data[data$patient_id %in% patients_chemo,]
```

```{r}
# run cox regression based on groups
fit.surv <- survfit(Surv(survival, status) ~ Factor8_group, data = data)

ggsurv <- ggsurvplot(fit.surv,
           data = data,
           risk.table = TRUE,
           conf.int = TRUE,
           size = 1,
           pval = TRUE,
           palette = c("#0000FF", "#FB0207"),
           legend.labs = c("t_lo", "t_hi"),
           ggtheme = theme_light(),
           tables.height = 0.23,
           xlim = c(0, 10),
           pval.coord = c(0,0),
           pval.size = 4.5,
           conf.int.alpha = 0.25)

ggsurv$plot <- ggsurv$plot + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text=element_text(family="Helvetica", size=12)) +xlab("Time (years)")+ 
  theme(strip.text.x = element_blank(),
      strip.background = element_rect(colour="white", fill="white"),
      legend.position=c(.85,.85))

ggsurv$table <- ggsurv$table + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                                      text=element_text(family="Helvetica", size=12))+ggtitle("")


pdf("results/figures/Survival/survival_curves_DDHM8_scores_txn_BeatAML_median.pdf", width = 4, height = 5)
print(ggsurv, newpage = FALSE)
dev.off()
```

## scores in sorafenib patients
```{r}
clinics_sorafenib <- clinics[grepl("Sorafenib", clinics$cumulativeTreatmentRegimens),]
nrow(clinics_sorafenib)
clinics_sorafenib <- clinics_sorafenib[grepl("FLT3", clinics_sorafenib$variantSummary) | clinics_sorafenib$FLT3.ITD == "positive",]
nrow(clinics_sorafenib)
```

## Extract transcriptomics
```{r}
txn <- gex[,colnames(gex) %in% c("stable_id", clinics_sorafenib$patient_id)]
txn$stable_id <- rownames(txn)
condition <- data.frame(patient_id = colnames(txn))
```

## Calculate scores transcriptomics
```{r}
for (i in 1:11) {
  fname <- paste0("Factor", i)
  myFactor1 <- RNAfac[RNAfac$factor==fname,]
  pick <- txn$stable_id %in% myFactor1$feature
  x=txn[pick,drop=FALSE,]
  x=t(x[,colnames(x) != "stable_id"])
  beta <- myFactor1$value[match(colnames(x), myFactor1$feature)]
  yval=t(t(x) * beta)
  yval=rowSums(yval)
  weight <- data.frame(yval)
  colnames(weight) <- fname
  weight$patient_id <- rownames(weight)
  yval <- data.frame(patient_id = names(yval), score = yval)
  yval$group <- as.integer(yval$score > median(yval$score))+1
  colnames(yval)[2:3] <- paste0(fname, "_", colnames(yval)[2:3])
  condition <- merge(condition, yval, by = "patient_id")
}
```

```{r}
clinical =clinics[,colnames(clinics) %in% c("patient_id","status", "survival")]
data <- merge(condition, clinical, by = "patient_id")
```

```{r}
colnames(data)[8]
```
```{r}
colnames(data)[8+14]
```

```{r}
a <- ggplot(data = data, mapping = aes(x = data[, 8], y =data[, 8+14])) +
          geom_point(size = 1) + xlab("DDHM4 (MOFA)") + ylab("DDHM4 (Transcriptomics score inverted sign)")+
          sm_statCorr(corr_method = "pearson", size = 0.1) +
  geom_smooth(method=lm, level=0.95, colour = "#BED797", fill = "#BED797", alpha=0.25)+
  theme_light()+ 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                                      text=element_text(family="Helvetica", size=12))
a
ggsave("results/figures/Associations/scatter_plots_MOFA_vs_transcriptomics_Factor_4_Sorafenib_v2_BeatAML.pdf", width = 4, height = 4, units = "in")
```


```{r}
# run cox regression based on groups
fit.surv <- survfit(Surv(survival, status) ~ Factor4_group, data = data)

ggsurv <- ggsurvplot(fit.surv,
           data = data,
           risk.table = TRUE,
           conf.int = TRUE,
           size = 1,
           pval = TRUE,
           palette = c("#0000FF", "#FB0207"),
           legend.labs = c("t_lo", "t_hi"),
           ggtheme = theme_light(),
           tables.height = 0.23,
           xlim = c(0, 5),
           pval.coord = c(0,0),
           pval.size = 4.5,
           conf.int.alpha = 0.25)

ggsurv$plot <- ggsurv$plot + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text=element_text(family="Helvetica", size=12)) +xlab("Time (years)")+ 
  theme(strip.text.x = element_blank(),
      strip.background = element_rect(colour="white", fill="white"),
      legend.position=c(.85,.85))

ggsurv$table <- ggsurv$table + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                                      text=element_text(family="Helvetica", size=12))+ggtitle("")


pdf("results/figures/Survival/survival_curves_DDHM4_scores_txn_BeatAML_median_sorafenib.pdf", width = 4, height = 5)
print(ggsurv, newpage = FALSE)
dev.off()
```



```{r}
clinics_venetoclax_azacitidine <- clinics[grepl("Venetoclax", clinics$currentRegimen),]
clinics_venetoclax_azacitidine <- clinics_venetoclax_azacitidine[grepl("Azacitidine", clinics_venetoclax_azacitidine$currentRegimen),]
nrow(clinics_venetoclax_azacitidine)
```

## Extract transcriptomics
```{r}
txn <- gex[,colnames(gex) %in% c("stable_id", clinics_venetoclax_azacitidine$patient_id)]
txn$stable_id <- rownames(txn)
condition <- data.frame(patient_id = colnames(txn))
```

## Calculate scores transcriptomics
```{r}
for (i in 1:11) {
  fname <- paste0("Factor", i)
  myFactor1 <- RNAfac[RNAfac$factor==fname,]
  pick <- txn$stable_id %in% myFactor1$feature
  x=txn[pick,drop=FALSE,]
  x=t(x[,colnames(x) != "stable_id"])
  beta <- myFactor1$value[match(colnames(x), myFactor1$feature)]
  yval=t(t(x) * beta)
  yval=rowSums(yval)
  weight <- data.frame(yval)
  colnames(weight) <- fname
  weight$patient_id <- rownames(weight)
  yval <- data.frame(patient_id = names(yval), score = yval)
  yval$group <- as.integer(yval$score > median(yval$score))+1
  colnames(yval)[2:3] <- paste0(fname, "_", colnames(yval)[2:3])
  condition <- merge(condition, yval, by = "patient_id")
}
```

```{r}
clinical =clinics[,colnames(clinics) %in% c("patient_id","status", "survival")]
data <- merge(condition, clinical, by = "patient_id")
```

```{r}
# run cox regression based on groups
fit.surv <- survfit(Surv(survival, status) ~ Factor2_group, data = data)

ggsurv <- ggsurvplot(fit.surv,
           data = data,
           risk.table = TRUE,
           conf.int = TRUE,
           size = 1,
           pval = TRUE,
           palette = c("#0000FF", "#FB0207"),
           legend.labs = c("t_lo", "t_hi"),
           ggtheme = theme_light(),
           tables.height = 0.23,
           xlim = c(0, 2),
           pval.coord = c(0,0),
           pval.size = 4.5,
           conf.int.alpha = 0.25)

ggsurv$plot <- ggsurv$plot + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text=element_text(family="Helvetica", size=12)) +xlab("Time (years)")+ 
  theme(strip.text.x = element_blank(),
      strip.background = element_rect(colour="white", fill="white"),
      legend.position=c(.85,.85))

ggsurv$table <- ggsurv$table + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                                      text=element_text(family="Helvetica", size=12))+ggtitle("")


pdf("results/figures/Survival/survival_curves_DDHM2_scores_txn_BeatAML_median_venetoclax_azacitidine.pdf", width = 4, height = 5)
print(ggsurv, newpage = FALSE)
dev.off()
```
