---
title: "R Notebook"
output: html_notebook
---

Change sign figures :
1. Pharmacoproteomic cohort: DDHM2 (fig s3a ) x 

DDHM cohort:
2. DDHM8 (transcriptome only) fig s6a x
3. DDHM4 (transcriptome only) fig s6b x

Check sign single cell :
*The problem with the single cells data is that i can not see which ones has flipped and which ones that has not without more information e.g, expression of top positive features showing the same pattern as the projected DDHM score. This may mean that the cell type association with the single cell data may be the reversed


### clean environment
```{r}
rm(list=ls())
```

### set directory
```{r setup}
    knitr::opts_knit$set(root.dir = normalizePath("/home/flomik/Code/MOFA_AML/")) 
```

```{r}
library("readxl")
library("data.table")
library(dplyr)
```

## extract top feature per factor
```{r}
RNA_factors=NULL
for (i in 1:11){
  factorName=paste0("Factor",i,"_RNAseq")
  cat(factorName," ")
  dat=read_excel("/home/flomik/Code/Hallmarks_AML/data/Ngias/AML_MOFAplus_top_factors_v04_f600 (1).xlsx",sheet=factorName)
  dat=data.table(dat)
  print(nrow(dat))
  RNA_factors[[i]]=dat
}

RNAfac=do.call(rbind,RNA_factors)
RNAfac$feature=gsub(".g","",RNAfac$feature)

## set thresholds for each factor
factorNames=unique(RNAfac$factor)
topGeneThres=c(9,10,2,6,4,1,2,3,3,6,3)
names(topGeneThres)=factorNames
```

```{r}
scores <- read.csv("/home/flomik/Code/MOFA_AML/results/scores_BeatAML_MOFA_transcriptomics.csv")
```

## transcriptomics
```{r}
gex <- read.delim("/home/flomik/Data/external_data_sets/BeatAML/beataml_waves1to4_norm_exp_dbgap.txt",  row.names = NULL)
gex$display_label <- NULL
rownames(gex) <- gex$stable_id
gex$stable_id <- NULL
```

### matching between scores and gene expression
```{r}
mypatients=intersect(scores$patient_id, colnames(gex))
length(mypatients)
gex=gex[,c(match(mypatients,colnames(gex)))]
sum(scores$patient_id!=colnames(gex)) #check if they are the same order
```

```{r}
col <- c("#584588",
"#4076A9",
"#ABCADA",
"#BED797",
"#5E9F48",
"#F1EFB3",
"#EEC17F",
"#E88840",
"#E99D9B",
"#CE3D31",
"#BCB2C9")
```

BeatAML -> 4, 7, 8

```{r}
#scores$Factor4_score <- -scores$Factor4_score
#scores$Factor7_score <- -scores$Factor7_score
#scores$Factor8_score <- -scores$Factor8_score

write.csv(scores,"/home/flomik/Code/MOFA_AML/results/scores_BeatAML_MOFA_transcriptomics_scores_inverted_4_7_8.csv")
```

```{r}
list_plots <- list()

for (i in 1:11){
  score_name <- paste0("Factor", i,"_score")
  score <- scores[, colnames(scores) == score_name]
  sheet <- RNA_factors[[i]]
  #pos <- sheet[sheet$sign == "+",]
  pos <- sheet
  pos <- pos[order(pos$value, decreasing = TRUE),]
  top_feat <- gsub(".g", "", pos$feature[1])
  top_name <- pos$label[1]
  gene <- data.frame(t(gex[rownames(gex) == top_feat,]))[,1]
  sign <- pos$sign[1]
  
  a <- ggplot(data = scores, mapping = aes(x = gene, y = score)) +
          geom_point(size = 1) + xlab(top_name) + 
      ylab(paste0("DDHM", i, " (Transcriptomics score)"))+
          sm_statCorr(corr_method = "pearson", size = 0.1) + 
      geom_smooth(method=lm, level=0.95, colour = col[i], fill = col[i], alpha=0.25)+ theme_light()+ 
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text=element_text(family="Helvetica", size=12))
  
  ggsave(paste0("results/figures/sign/beatAML/scatter_plot_top_value_Factor_", i, ".pdf"))
  print(a)
  list_plots[[i]] <- a
}
```
