---
title: "R Notebook"
output: html_notebook
---

1. Compare ELBO (Evidence Lower Bound) == DONE

2. Run alternative MOFA models trained on random subsets of decreasing sample sizes (i.e., 90%, 80%, 70%, 60%, and 50%). preserving omics proportions. Each threshold was repeated 5 times with different random seed.

3. 50-50 % split
Train non-overlapping donor sets, training independent MOFA models on each, and correlating the feature weights across matching factors (also 15 random splits). Results were similar to those from 50% vs. the full model, suggesting that the conserved factors are not driven by individual-specific data 

4. samples with increasing numbers of omics layers (i.e., subsets with at least 2, 3, or 4 omics layers)

list_datasets[[i]] <- data_sub

### clean environment
```{r}
rm(list=ls())
```

### set directory (replace by your repository)
```{r setup}
    knitr::opts_knit$set(root.dir = normalizePath("/home/flomik/Code/MOFA_AML/")) 
```


```{r}
library(MultiAssayExperiment)
library(MOFA2)
library(dplyr)
```

## load data
```{r}
load("/home/flomik/Data/MOFA_R_Objects/MOFA_v04_train_list.RData")
mofa.sel <- select_model(mofa.list)


data_2 <- list(
  Methylation = mofa.sel@data$DNAm$AML, 
  DSRT = mofa.sel@data$DSRT$AML,
  Mutations =mofa.sel@data$Mutations$AML,
  Proteins = mofa.sel@data$Proteomics$AML,
  Genes = mofa.sel@data$RNAseq$AML
  )

colnames(data_2$DSRT) == colnames(data_2$Genes)
```

```{r}
a <-colSums(is.na(data_2$DSRT)) != nrow(data_2$DSRT)
b <-colSums(is.na(data_2$Genes)) != nrow(data_2$Genes)
c <-colSums(is.na(data_2$Mutations)) != nrow(data_2$Mutations)
d <-colSums(is.na(data_2$Methylation)) != nrow(data_2$Methylation)
e <-colSums(is.na(data_2$Proteins)) != nrow(data_2$Proteins)

table_missing <- rbind(a, b, c, d, e)
table_missing <- 1*table_missing

rownames(table_missing) <- c("DSRT", "Genes", "Mutations", "Methylation", "Proteins")

table_missing <- data.frame(table_missing)

table_missing$percentage <- as.vector(rowSums(table_missing)/118)
```

```{r}
per <- c(0.5)
name <- c("50")

x <- sample(colnames(table_missing)[table_missing[1,] == 1], length(colnames(table_missing)[table_missing[1,] == 1])*per)
length(x)
x_2 <- colnames(table_missing)[table_missing[1,] == 1 & !colnames(table_missing) %in% x]
length(x_2)

y <- sample(colnames(table_missing)[table_missing[1,] == 0], length(colnames(table_missing)[table_missing[1,] == 0])*per)
length(y)
y_2 <- colnames(table_missing)[table_missing[1,] == 0 & !colnames(table_missing) %in% y]
length(y_2)

table_missing_2 <- table_missing[, colnames(table_missing) %in% c(x, y)]

data_sub_1 <- list(
    DSRT = mofa.sel@data$DSRT$AML[, colnames(mofa.sel@data$DSRT$AML) %in% colnames(table_missing_2)],
    Genes = mofa.sel@data$RNAseq$AML[, colnames(mofa.sel@data$RNAseq$AML) %in% colnames(table_missing_2)],
    Mutations =mofa.sel@data$Mutations$AML[, colnames(mofa.sel@data$Mutations$AML) %in% colnames(table_missing_2)],
    Methylation = mofa.sel@data$DNAm$AML[, colnames(mofa.sel@data$DNAm$AML) %in% colnames(table_missing_2)], 
    Proteins = mofa.sel@data$Proteomics$AML[, colnames(mofa.sel@data$Proteomics$AML) %in% colnames(table_missing_2)])
  
   colnames(data_sub_1$DSRT) == colnames(data_sub_1$Genes)
   saveRDS(data_sub_1, paste0("/home/flomik/Data/MOFA_R_Objects/Robustness/v1_independant_50_data_input_MOFA.rds"))
   
   
table_missing_3 <- table_missing[, colnames(table_missing) %in% c(x_2, y_2)]

data_sub_2 <- list(
    DSRT = mofa.sel@data$DSRT$AML[, colnames(mofa.sel@data$DSRT$AML) %in% colnames(table_missing_3)],
    Genes = mofa.sel@data$RNAseq$AML[, colnames(mofa.sel@data$RNAseq$AML) %in% colnames(table_missing_3)],
    Mutations =mofa.sel@data$Mutations$AML[, colnames(mofa.sel@data$Mutations$AML) %in% colnames(table_missing_3)],
    Methylation = mofa.sel@data$DNAm$AML[, colnames(mofa.sel@data$DNAm$AML) %in% colnames(table_missing_3)], 
    Proteins = mofa.sel@data$Proteomics$AML[, colnames(mofa.sel@data$Proteomics$AML) %in% colnames(table_missing_3)])
  
   colnames(data_sub_2$DSRT) == colnames(data_sub_2$Genes)
   saveRDS(data_sub_2, paste0("/home/flomik/Data/MOFA_R_Objects/Robustness/v2_independant_50_data_input_MOFA.rds"))
  
colnames(data_sub_1$DSRT) == colnames(data_sub_2$DSRT)
```

```{r}
list_datasets <- list()
list_datasets[[1]] <- data_sub_1
list_datasets[[2]] <- data_sub_2

name <- c("v1_independant_50","v2_independant_50", "Methylation", "DSRT", "Mutations","Proteins", "Genes", ">5",">4")
```

## remove 1 omics layer one by one
```{r}
for (i in 3:7) {
  name_ana <- name[i]
  print(name_ana)
  data_3 <- data_2
  data_3[[name_ana]][!is.na(data_3[[name_ana]])] <- NA
  list_datasets[[i]] <- data_3
}
```

## increasing numbers of omics layers (i.e., subsets with at least 2, 3, or 4 omics layers), and by completely removing one omics layer (i.e., setting ‘NA’ to all features of the removed layer)
```{r}
Methylation_names = colnames(mofa.sel@data$DNAm$AML)[colSums(is.na(mofa.sel@data$DNAm$AML)) < 5000]
Genes_names = colnames(mofa.sel@data$RNAseq$AML)[colSums(is.na(mofa.sel@data$RNAseq$AML)) < 5000]
Proteins_names = colnames(mofa.sel@data$Proteomics$AML)[colSums(is.na(mofa.sel@data$Proteomics$AML)) < 1]
Mutations_names = colnames(mofa.sel@data$Mutations$AML)[colSums(is.na(mofa.sel@data$Mutations$AML)) < 1]
DSRT_names = colnames(mofa.sel@data$DSRT$AML)[colSums(is.na(mofa.sel@data$DSRT$AML)) < 528]

data_3 <- data_2
data_3$Methylation <- data_3$Methylation[ ,colnames(data_3$Methylation) %in% DSRT_names]
data_3$DSRT <- data_3$DSRT[ ,colnames(data_3$DSRT) %in% DSRT_names]
data_3$Genes <- data_3$Genes[,colnames(data_3$Genes) %in% DSRT_names]
data_3$Mutations <- data_3$Mutations[ ,colnames(data_3$Mutations) %in% DSRT_names]
data_3$Proteins <- data_3$Proteins[ ,colnames(data_3$Proteins) %in% DSRT_names]

list_datasets[[8]] <- data_3

data_3 <- data_2
data_3$Methylation <- data_3$Methylation[ ,colnames(data_3$Methylation) %in% Methylation_names]
data_3$DSRT <- data_3$DSRT[ ,colnames(data_3$DSRT) %in% Methylation_names]
data_3$Genes <- data_3$Genes[,colnames(data_3$Genes) %in% Methylation_names]
data_3$Mutations <- data_3$Mutations[ ,colnames(data_3$Mutations) %in% Methylation_names]
data_3$Proteins <- data_3$Proteins[ ,colnames(data_3$Proteins) %in% Methylation_names]

list_datasets[[9]] <- data_3

```

```{r}
#' Run MOFA by francesco
#'
#' @param data MultiAssayExperiment object
#' @param name title of analysis
#' @return mofa object
#' @export
MOFA_fran <- function(data, name_ana){
  ## create model
  MOFAobject <- create_mofa(data)
  ## set parameters
  data_opts <- get_default_data_options(MOFAobject)
  data_opts$scale_views <- T
  model_opts <- get_default_model_options(MOFAobject)
  model_opts$num_factors <- 15
  model_opts$likelihoods <-  c("gaussian","gaussian","bernoulli","gaussian","gaussian")
  ## run model
  set.seed(12345)
  seeds <- floor(10^7*runif(5))

  mofa.list <- list()
  
  for(i in 1:5){
    # Training options
    # - maxiter: number of iterations
    # - convergence_mode: "fast", "medium", "slow". For exploration, the fast mode is good enough.
    # - startELBO: initial iteration to compute the ELBO (the objective function used to assess convergence)
    # - freqELBO: frequency of computations of the ELBO (the objective function used to assess convergence)
    # - drop_factor_threshold: minimum variance explained criteria to drop factors while training. Default is -1 (no dropping of factors)
    # - gpu_mode: use GPU mode? This needs cupy installed and a functional GPU, see https://cupy.chainer.org/
    # - verbose: verbose mode?
    # - seed: random seed
    train_opts <- get_default_training_options(MOFAobject)
    train_opts$convergence_mode <- "slow"
    train_opts$drop_factor_threshold <- 0.02
    train_opts$maxiter <- 10000
    train_opts$seed <- seeds[i]
    
    mofa.toTrain <- prepare_mofa(MOFAobject,
                               data_options = data_opts,
                               model_options = model_opts,
                               training_options = train_opts )
      outfile <- paste0("/home/flomik/Data/MOFA_R_Objects/Robustness/", name_ana, "_MOFA.hdf5")
      mofa.list[[i]] <- run_mofa(mofa.toTrain, outfile, use_basilisk = T)
      mofa.sel <- select_model(mofa.list)
    }
  #saveRDS(mofa.sel, file=paste0("/home/flomik/Data/MOFA_R_Objects/Robustness/", name, "_MOFA.rds"))
  return(mofa.sel)
}
```


```{r}
for (j in 3:9) {
  data <- MOFA_fran(list_datasets[[j]], name[[j]])
}
```

