---
title: "R Notebook"
output: html_notebook
---

1. Compare ELBO (Evidence Lower Bound) == DONE

2. Run alternative MOFA models trained on random subsets of decreasing sample sizes (i.e., 90%, 80%, 70%, 60%, and 50%). preserving omics proportions. Each threshold was repeated 5 times with different random seed.

3. 50-50 % split
Train non-overlapping donor sets, training independent MOFA models on each, and correlating the feature weights across matching factors (also 15 random splits). Results were similar to those from 50% vs. the full model, suggesting that the conserved factors are not driven by individual-specific data 

4. samples with increasing numbers of omics layers (i.e., subsets with at least 2, 3, or 4 omics layers)

list_datasets[[i]] <- data_sub

### clean environment
```{r}
rm(list=ls())
```

### set directory (replace by your repository)
```{r setup}
    knitr::opts_knit$set(root.dir = normalizePath("/home/flomik/Code/MOFA_AML/")) 
```


```{r}
library(MultiAssayExperiment)
library(MOFA2)
library(dplyr)
```

## load data
```{r}
load("/home/flomik/Data/MOFA_R_Objects/MOFA_v04_train_list.RData")
mofa.sel <- select_model(mofa.list)
```

```{r}
name <- c("v1_independant_50","v2_independant_50", "Methylation", "DSRT", "Mutations","Proteins", "Genes", ">5",">4")
```

```{r}
weights_main <- get_weights(mofa.sel, views = "all", factors = "all")
weights_main_2 <- rbind(weights_main$DNAm, weights_main$DSRT, weights_main$Mutations, weights_main$Proteomics, weights_main$RNAseq)
```

```{r}
for (i in 1:9) {
  filepath <- paste0("/home/flomik/Data/MOFA_R_Objects/Robustness/", name[i], "_MOFA.hdf5")
  model_comp <- load_model(filepath)
  weights_comp <- get_weights(model_comp, views = "all", factors = "all")
  weights_comp_2 <- rbind(weights_comp$Methylation, weights_comp$DSRT, weights_comp$Mutations, weights_comp$Proteins, weights_comp$Genes)
  print(name[i])
  print(ncol(weights_comp_2))
  correlation <- cor(weights_main_2, weights_comp_2, method = 'pearson')
  max_values <- c(colMaxs(abs(correlation)), -colMaxs(abs(correlation)))
  correlation[!correlation %in% c(max_values)] <- 0

  correlation_2 <- melt(correlation)

  correlation_2 <- correlation_2[correlation_2$value != 0,]
  colnames(correlation_2)[1:2] <- c("main", "comparison")
  correlation_2$value <- abs(correlation_2$value)
  correlation_2 <- correlation_2[order(correlation_2$value),]
  correlation_2$comparison <- factor(correlation_2$comparison, levels = correlation_2$comparison)
  
  p<-ggplot(data=correlation_2, aes(x=value, y=comparison)) +
  geom_bar(stat="identity")+ theme_classic()+xlab("Features weights correlation coefficient")+ylab("")
  p
  ggsave(paste0("results/figures/MOFA/correlation_plots_MOFA_", name[i],".pdf"))
}
```


1.extract weights per sample
2. find matched factors between main model and submodel = best-matching approach based on correlation of feature loadings
3. calculate coefficient correlation (Pearson for each factor and each view)
4. Make plot

```{r}
correlation <- cor(weights_main_2, weights_comp_2, method = 'pearson')
max_values <- c(colMaxs(abs(correlation)), -colMaxs(abs(correlation)))
#correlation_test <- correlation
#correlation_test[correlation_test %in% max_values] <- 0
#max_values_2 <-colMaxs(correlation_test)

#rownames(weights_main$DSRT) == rownames(weights_comp$DSRT)
#correlation[!correlation %in% c(max_values, max_values_2)] <- 0
correlation[!correlation %in% c(max_values)] <- 0

correlation_2 <- melt(correlation)

correlation_2 <- correlation_2[correlation_2$value != 0,]
colnames(correlation_2)[1:2] <- c("main", "comparison")
correlation_2$value <- abs(correlation_2$value)
correlation_2 <- correlation_2[order(correlation_2$value),]
correlation_2$comparison <- factor(correlation_2$comparison, levels = correlation_2$comparison)
```

```{r}
library(ggplot2)

p<-ggplot(data=correlation_2, aes(x=value, y=comparison)) +
  geom_bar(stat="identity")
p
   
```


