---
title: "R Notebook"
output: html_notebook
---

1. Compare ELBO (Evidence Lower Bound) == DONE

2. Run alternative MOFA models trained on random subsets of decreasing sample sizes (i.e., 90%, 80%, 70%, 60%, and 50%). preserving omics proportions. Each threshold was repeated 5 times with different random seed.

3. 50-50 % split
Train non-overlapping donor sets, training independent MOFA models on each, and correlating the feature weights across matching factors (also 15 random splits). Results were similar to those from 50% vs. the full model, suggesting that the conserved factors are not driven by individual-specific data 

4. samples with increasing numbers of omics layers (i.e., subsets with at least 2, 3, or 4 omics layers)

list_datasets[[i]] <- data_sub

### clean environment
```{r}
rm(list=ls())
```

### set directory (replace by your repository)
```{r setup}
    knitr::opts_knit$set(root.dir = normalizePath("/home/flomik/Code/MOFA_AML/")) 
```


```{r}
library(MultiAssayExperiment)
library(MOFA2)
library(dplyr)
```

## load data
```{r}
load("/home/flomik/Data/MOFA_R_Objects/MOFA_v04_train_list.RData")
mofa.sel <- select_model(mofa.list)


data_2 <- list(
  DSRT = mofa.sel@data$DSRT$AML,
  Genes = mofa.sel@data$RNAseq$AML,
  Mutations =mofa.sel@data$Mutations$AML,
  Methylation = mofa.sel@data$DNAm$AML, 
  Proteins = mofa.sel@data$Proteomics$AML)

colnames(data_2$DSRT) == colnames(data_2$Genes)
```

```{r}
a <-colSums(is.na(data_2$DSRT)) != nrow(data_2$DSRT)
b <-colSums(is.na(data_2$Genes)) != nrow(data_2$Genes)
c <-colSums(is.na(data_2$Mutations)) != nrow(data_2$Mutations)
d <-colSums(is.na(data_2$Methylation)) != nrow(data_2$Methylation)
e <-colSums(is.na(data_2$Proteins)) != nrow(data_2$Proteins)

table_missing <- rbind(a, b, c, d, e)
table_missing <- 1*table_missing

rownames(table_missing) <- c("DSRT", "Genes", "Mutations", "Methylation", "Proteins")

table_missing <- data.frame(table_missing)

table_missing$percentage <- as.vector(rowSums(table_missing)/118)
```

```{r}
percent <- c(0.9, 0.8, 0.7, 0.6, 0.5)
name <- c("90", "80", "70", "60", "50")
```

```{r}
list_datasets <- list()
list_models <- list()

for (i in 1:5) {
  per <- percent[i]
  name_ana <- name[i]
  print(c(name_ana, per))
  
  x <- sample(colnames(table_missing)[table_missing[1,] == 1], length(colnames(table_missing)[table_missing[1,] == 1])*per)
  length(x)
  y <- sample(colnames(table_missing)[table_missing[1,] == 0], length(colnames(table_missing)[table_missing[1,] == 0])*per)
  length(y)

  table_missing_2 <- table_missing[, colnames(table_missing) %in% c(x, y)]
  table_missing_2$percentage <- as.vector(rowSums(table_missing_2)/ncol(table_missing_2))
  print(table_missing_2$percentage)
  print(table_missing$percentage)
  print(dim(table_missing_2))

  data_sub <- list(
    DSRT = mofa.sel@data$DSRT$AML[, colnames(mofa.sel@data$DSRT$AML) %in% colnames(table_missing_2)],
    Genes = mofa.sel@data$RNAseq$AML[, colnames(mofa.sel@data$RNAseq$AML) %in% colnames(table_missing_2)],
    Mutations =mofa.sel@data$Mutations$AML[, colnames(mofa.sel@data$Mutations$AML) %in% colnames(table_missing_2)],
    Methylation = mofa.sel@data$DNAm$AML[, colnames(mofa.sel@data$DNAm$AML) %in% colnames(table_missing_2)], 
    Proteins = mofa.sel@data$Proteomics$AML[, colnames(mofa.sel@data$Proteomics$AML) %in% colnames(table_missing_2)])
  
   colnames(data_sub$DSRT) == colnames(data_sub$Genes)
   saveRDS(data_sub, paste0("/home/flomik/Data/MOFA_R_Objects/Robustness/", name_ana, "_data_input_MOFA.rds"))
   list_datasets[[i]] <- data_sub
}
```

```{r}
#' Run MOFA by francesco
#'
#' @param data MultiAssayExperiment object
#' @param name title of analysis
#' @return mofa object
#' @export
MOFA_fran <- function(data, name){
  ## create model
  MOFAobject <- create_mofa(data)
  ## set parameters
  data_opts <- get_default_data_options(MOFAobject)
  data_opts$scale_views <- T
  model_opts <- get_default_model_options(MOFAobject)
  model_opts$num_factors <- 15
  model_opts$likelihoods <-  c("gaussian","gaussian","bernoulli","gaussian","gaussian")
  ## run model
  set.seed(12345)
  seeds <- floor(10^7*runif(5))

  mofa.list <- list()
  
  for(i in 1:5){
    # Training options
    # - maxiter: number of iterations
    # - convergence_mode: "fast", "medium", "slow". For exploration, the fast mode is good enough.
    # - startELBO: initial iteration to compute the ELBO (the objective function used to assess convergence)
    # - freqELBO: frequency of computations of the ELBO (the objective function used to assess convergence)
    # - drop_factor_threshold: minimum variance explained criteria to drop factors while training. Default is -1 (no dropping of factors)
    # - gpu_mode: use GPU mode? This needs cupy installed and a functional GPU, see https://cupy.chainer.org/
    # - verbose: verbose mode?
    # - seed: random seed
    train_opts <- get_default_training_options(MOFAobject)
    train_opts$convergence_mode <- "slow"
    train_opts$drop_factor_threshold <- 0.02
    train_opts$maxiter <- 10000
    train_opts$seed <- seeds[i]
    
    mofa.toTrain <- prepare_mofa(MOFAobject,
                               data_options = data_opts,
                               model_options = model_opts,
                               training_options = train_opts )
      outfile <- paste0("/home/flomik/Data/MOFA_R_Objects/Robustness/", name, "_MOFA.rds")
      mofa.list[[i]] <- run_mofa(mofa.toTrain, outfile, use_basilisk = T)
      mofa.sel <- select_model(mofa.list)
    }
  #saveRDS(mofa.sel, file=paste0("/home/flomik/Data/MOFA_R_Objects/Robustness/", name, "_MOFA.rds"))
  return(mofa.sel)
}
```


```{r}
for (j in 1:5) {
  data <- MOFA_fran(list_datasets[[j]], name[[j]])
  list_models[[i]] <- data
}
```



## extract factors
```{r}
mofa_sub <- readRDS("/home/flomik/Data/MOFA_R_Objects/Robustness/90_data_input_MOFA.rds")
```

1.extract weights per sample
2. find matched factors between main model and submodel
3. calculate coefficient correlation (Pearson for each factor)
4. Make plot

## calculate mean features weight
```{r}
weights <- get_weights(mofa_sub)
```

