---
title: "survival MOFA Factor 8"
output: html_notebook
---


Analysis:



### clean environment
```{r}
rm(list=ls())
```

### set directory

```{r setup}
    knitr::opts_knit$set(root.dir = normalizePath("/home/flomik/Code/MOFA_AML")) 
```

```{r}
library(MOFA2)
library(data.table)
library(readxl)
library(ggplot2)
library(survival)
library(survminer)
library("smplot2")
library(MultiAssayExperiment)
library(gridExtra)
library(xlsx)
```

## load data
```{r}
load("/home/flomik/Data/MOFA_R_Objects/AML_MAexp_v05.RData")
ls()
source("/home/flomik/Code/Hallmarks_AML/src/functions_Hallmarks.R")
annot.df <- read.table("/home/flomik/Data/MOFA_R_Objects/MOFA_v04_annot_features.txt", sep="\t", header = T, stringsAsFactors = F)
load("/home/flomik/Data/MOFA_R_Objects/MOFA_v04_train_list.RData")
```

## Select MOFA object
```{r}
mofa.sel <- select_model(mofa.list)
```

## extract top features
```{r}
path_factors <- "/home/flomik/Code/MOFA_AML/data/Factors/AML_MOFAplus_top_factors_v04_f600 (1).xlsx"
RNA_factors=NULL
for (i in 1:11){
  factorName=paste0("Factor",i,"_RNAseq")
  cat(factorName," ")
  dat=read_excel(path_factors,sheet=factorName)
  dat=data.table(dat)
  print(nrow(dat))
  RNA_factors[[i]]=dat
}
RNAfac=do.call(rbind,RNA_factors)
RNAfac$feature=gsub(".g","",RNAfac$feature)
## set thresholds for each factor
factorNames=unique(RNAfac$factor)
topGeneThres=c(9,10,2,6,4,1,2,3,3,6,3)
names(topGeneThres)=factorNames
```

## Extract transcriptomics
```{r}
gex <- data.frame(mofa.sel@data[[5]]$AML)
gex <-gex[!is.na(gex),]
rownames(gex) <- gsub(".g", "", rownames(gex))
gex$stable_id <- rownames(gex)
txn <- gex[!is.na(gex$ALBB13002),]
txn$ALG200931 <- NULL
condition <- data.frame(patient_id = colnames(txn))
```

## Calculate factor scores transcriptomics
```{r}
for (i in 1:11) {
  fname <- paste0("Factor", i)
  myFactor1 <- RNAfac[RNAfac$factor==fname,]
  pick <- txn$stable_id %in% myFactor1$feature
  x=txn[pick,drop=FALSE,]
  x=t(x[,colnames(x) != "stable_id"])
  beta <- myFactor1$value[match(colnames(x), myFactor1$feature)]
  yval=t(t(x) * beta)
  yval=rowSums(yval)
  weight <- data.frame(yval)
  colnames(weight) <- fname
  weight$patient_id <- rownames(weight)
  yval <- data.frame(patient_id = names(yval), score = yval)
  yval$group <- as.integer(yval$score > median(yval$score))+1
  colnames(yval)[2:3] <- paste0(fname, "_", colnames(yval)[2:3])
  condition <- merge(condition, yval, by = "patient_id")
}
```

## extract DDHM values
```{r}
factors <- read_xlsx("/home/flomik/Code/MOFA_AML/data/Factors/DDHMpatValue.xlsx")
patients <- factors$patient_id
rownames(factors) <- factors$patient_id
factors$sample <- NULL
factors$patient_id <- NULL
factors <- data.frame(lapply(factors, as.numeric))
factors$patient_id <- patients
txn <- condition[,colnames(condition) %in% c("patient_id", colnames(condition)[grepl("score", colnames(condition))])]
data <- merge(factors, txn, by = "patient_id")
```

```{r}
myplots <- vector('list', 11)

for (i in 2:12) {
  myplots[[i-1]] <- local({
        i <- i
        a <- ggplot(data = data, mapping = aes(x = data[, i], y =data[, i+11])) +
          geom_point(size = 1) + xlab("") + ylab("")+theme_classic()+
          sm_statCorr( color = "#0f993d", corr_method = "spearman", linetype = "dashed")+
          ggtitle(paste0("DDHM", i-1))
        print(a)
    })
}
```

```{r}
a <- ggplot(data = data, mapping = aes(x = data[, 9], y =data[, 9+11])) +
          geom_point(size = 1) + xlab("DDHM8 (MOFA)") + ylab("DDHM8 (Transcriptomics score)")+
          sm_statCorr(corr_method = "pearson", size = 0.1) +
  geom_smooth(method=lm, level=0.95, colour = "#FF7F00", fill = "#FF7F00", alpha=0.25)+
  theme_light()+ 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                                      text=element_text(family="Helvetica", size=12))
a
ggsave("results/figures/scatter_plots_MOFA_vs_transcriptomics_Factor_8.pdf", width = 4, height = 4, units = "in")
```

```{r}
plots <- marrangeGrob(myplots, nrow = 4, ncol = 3, layout_matrix = matrix(
    seq_len(4 * 3),
    nrow = 4, ncol = 3,
    byrow = TRUE
  ))
ggsave("results/figures/scatter_plots_MOFA_vs_transcriptomics.pdf", plots, width = 8, height = 8, units = "in")
```

## load clinical data
```{r}
clinical_data <- data.frame(pData)
clinical_data <- clinical_data[, colnames(clinical_data) %in% c("time_days", "status", "patient_id")]
clinical_data$survival <- clinical_data$time_days/365
data <- merge(factors, condition, by = "patient_id")
data <- merge(clinical_data, data, by = "patient_id")
```


## separate groups between high Factor 8 and low Factor 8
```{r}
data$Factor8_group <- paste0("G", data$Factor8_group)
data$status
data$DDHM8_group <- NA
data$DDHM8_group[data$DDHM8 > median(data$DDHM8)] <- 2
data$DDHM8_group[data$DDHM8 < median(data$DDHM8)] <- 1
```

## Survival analysis
```{r}
# run cox regression based on groups
fit.surv <- survfit(Surv(survival, status) ~ DDHM8_group, data = data)

ggsurv <- ggsurvplot(fit.surv,
           data = data,
           risk.table = TRUE,
           conf.int = TRUE,
           size = 1,
           pval = TRUE,
           palette = c("#0000FF", "#FB0207"),
           legend.labs = c("m_lo", "m_hi"),
           ggtheme = theme_light(),
           tables.height = 0.23,
           xlim = c(0, 15),
           pval.coord = c(0,0),
           pval.size = 4.5,
           conf.int.alpha = 0.25)

ggsurv$plot <- ggsurv$plot + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text=element_text(family="Helvetica", size=12)) +xlab("Time (years)")+ 
  theme(strip.text.x = element_blank(),
      strip.background = element_rect(colour="white", fill="white"),
      legend.position=c(.85,.85))+ggtitle("")

ggsurv$table <- ggsurv$table + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                                      text=element_text(family="Helvetica", size=12))

ggsurv 
pdf("results/figures/survival_curves_DDHM8_median.pdf", width = 4, height = 5)
print(ggsurv, newpage = FALSE)
dev.off()
```

```{r}
# run cox regression based on groups
fit.surv <- survfit(Surv(survival, status) ~ Factor8_group, data = data)

ggsurv <- ggsurvplot(fit.surv,
           data = data,
           risk.table = TRUE,
           conf.int = TRUE,
           size = 1,
           pval = TRUE,
           palette = c("#0000FF", "#FB0207"),
           legend.labs = c("t_lo", "t_hi"),
           ggtheme = theme_light(),
           tables.height = 0.23,
           xlim = c(0, 15),
           pval.coord = c(0,0),
           pval.size = 4.5,
           conf.int.alpha = 0.25)

ggsurv$plot <- ggsurv$plot + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text=element_text(family="Helvetica", size=12)) +xlab("Time (years)")+ 
  theme(strip.text.x = element_blank(),
      strip.background = element_rect(colour="white", fill="white"),
      legend.position=c(.85,.85))

ggsurv$table <- ggsurv$table + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                                      text=element_text(family="Helvetica", size=12))+ggtitle("")

ggsurv
pdf("results/figures/survival_curves_DDHM8_scores_txn_KI_median.pdf", width = 4, height = 5)
print(ggsurv, newpage = FALSE)
dev.off()
```

## Survival DDHM8 BeatAML

### paths files / analysis
```{r}
name_ana <- "BeatAML"
path_txn <- "/home/flomik/Data/BeatAML/beataml_waves1to4_norm_exp_dbgap.txt"
path_clinical <- "/home/flomik/Code/Hallmarks_AML/data/Beta_AML/Publications/beataml_wv1to4_clinical.xlsx"
path_factors <- "/home/flomik/Code/MOFA_AML/data/Factors/AML_MOFAplus_top_factors_v04_f600 (1).xlsx"
```

## add categories
```{r}
categories_drug <- read.xlsx("/home/flomik/Code/MOFA_AML/data/Drugs/Annotation_curated_DrugBank_ChEMBL_class_revised_2021-10-28.xlsx", 1)
```

### import gene expression
```{r}
gex <- read.delim(path_txn,  row.names = NULL)
gex$display_label <- NULL
rownames(gex) <- gex$stable_id
gex$stable_id <- NULL
```

### import clinical data and preprocessing (rename columns, remove samples, harmonize, make sub-groups) / different for each cohort
```{r}
clinics <- read.xlsx(path_clinical, 1)
colnames(clinics)[colnames(clinics) == "vitalStatus"] <- "status"
colnames(clinics)[colnames(clinics) == "overallSurvival"] <- "follow_up_days"
colnames(clinics)[colnames(clinics) == "dbgap_rnaseq_sample"] <- "patient_id"

table(clinics$status)
## remove patients with unknow status
clinics <- clinics[clinics$status != "Unknown",]
clinics$status=ifelse(clinics$status=="Dead", 1, 0)

## calculate survival
clinics$survival <- clinics$follow_up_days/365
clinics <- clinics[!duplicated(clinics$patient_id),]
clinics <- clinics[!is.na(clinics$patient_id),]
rownames(clinics) <- clinics$patient_id

clinics$ELN2017_char <- clinics$ELN2017
clinics$ELN2017[!clinics$ELN2017_char %in% c("Adverse", "Favorable", "Intermediate")] <- NA

clinics$ELN2017[clinics$ELN2017 == "Favorable" & !is.na(clinics$ELN2017)] <- 1
clinics$ELN2017[clinics$ELN2017 == "Intermediate" & !is.na(clinics$ELN2017)] <- 2
clinics$ELN2017[clinics$ELN2017 == "Adverse" & !is.na(clinics$ELN2017)] <- 3
clinics$ELN2017 <- as.numeric(clinics$ELN2017)
table(clinics$ELN2017)
```

### matching between clinics and gene expression
```{r}
mypatients=intersect(clinics$patient_id, colnames(gex))
length(mypatients)
gex=gex[,c(match(mypatients,colnames(gex)))]
clinics=clinics[match(mypatients,clinics$patient_id),]
sum(clinics$patient_id!=colnames(gex)) #check if they are the same order
```

## Extract transcriptomics
```{r}
txn <- gex
txn$stable_id <- rownames(txn)
condition <- data.frame(patient_id = colnames(txn))
```

## Calculate scores transcriptomics
```{r}
for (i in 1:11) {
  fname <- paste0("Factor", i)
  myFactor1 <- RNAfac[RNAfac$factor==fname,]
  pick <- txn$stable_id %in% myFactor1$feature
  x=txn[pick,drop=FALSE,]
  x=t(x[,colnames(x) != "stable_id"])
  beta <- myFactor1$value[match(colnames(x), myFactor1$feature)]
  yval=t(t(x) * beta)
  yval=rowSums(yval)
  weight <- data.frame(yval)
  colnames(weight) <- fname
  weight$patient_id <- rownames(weight)
  yval <- data.frame(patient_id = names(yval), score = yval)
  yval$group <- as.integer(yval$score > median(yval$score))+1
  colnames(yval)[2:3] <- paste0(fname, "_", colnames(yval)[2:3])
  condition <- merge(condition, yval, by = "patient_id")
}
```


```{r}
clinical =clinics[,colnames(clinics) %in% c("patient_id","status", "survival")]
patients_chemo <- clinics$patient_id[grepl("Standard Chemotherapy", clinics$typeInductionTx)]
data <- merge(condition, clinical, by = "patient_id")
data <- data[data$patient_id %in% patients_chemo,]
```

```{r}
# run cox regression based on groups
fit.surv <- survfit(Surv(survival, status) ~ Factor8_group, data = data)

ggsurv <- ggsurvplot(fit.surv,
           data = data,
           risk.table = TRUE,
           conf.int = TRUE,
           size = 1,
           pval = TRUE,
           palette = c("#0000FF", "#FB0207"),
           legend.labs = c("t_lo", "t_hi"),
           ggtheme = theme_light(),
           tables.height = 0.23,
           xlim = c(0, 10),
           pval.coord = c(0,0),
           pval.size = 4.5,
           conf.int.alpha = 0.25)

ggsurv$plot <- ggsurv$plot + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text=element_text(family="Helvetica", size=12)) +xlab("Time (years)")+ 
  theme(strip.text.x = element_blank(),
      strip.background = element_rect(colour="white", fill="white"),
      legend.position=c(.85,.85))

ggsurv$table <- ggsurv$table + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                                      text=element_text(family="Helvetica", size=12))+ggtitle("")


pdf("results/figures/survival_curves_DDHM8_scores_txn_BeatAML_median.pdf", width = 4, height = 5)
print(ggsurv, newpage = FALSE)
dev.off()
```

## scores in venetoclax patients
```{r}
clinics_sorafenib <- clinics[grepl("Sorafenib", clinics$cumulativeTreatmentRegimens),]
nrow(clinics_sorafenib)
clinics_sorafenib <- clinics_sorafenib[grepl("FLT3", clinics_sorafenib$variantSummary) | clinics_sorafenib$FLT3.ITD == "positive",]
nrow(clinics_sorafenib)
```

## Extract transcriptomics
```{r}
txn <- gex[,colnames(gex) %in% c("stable_id", clinics_sorafenib$patient_id)]
txn$stable_id <- rownames(txn)
condition <- data.frame(patient_id = colnames(txn))
```

## Calculate scores transcriptomics
```{r}
for (i in 1:11) {
  fname <- paste0("Factor", i)
  myFactor1 <- RNAfac[RNAfac$factor==fname,]
  pick <- txn$stable_id %in% myFactor1$feature
  x=txn[pick,drop=FALSE,]
  x=t(x[,colnames(x) != "stable_id"])
  beta <- myFactor1$value[match(colnames(x), myFactor1$feature)]
  yval=t(t(x) * beta)
  yval=rowSums(yval)
  weight <- data.frame(yval)
  colnames(weight) <- fname
  weight$patient_id <- rownames(weight)
  yval <- data.frame(patient_id = names(yval), score = yval)
  yval$group <- as.integer(yval$score > median(yval$score))+1
  colnames(yval)[2:3] <- paste0(fname, "_", colnames(yval)[2:3])
  condition <- merge(condition, yval, by = "patient_id")
}
```

```{r}
clinical =clinics[,colnames(clinics) %in% c("patient_id","status", "survival")]
data <- merge(condition, clinical, by = "patient_id")
```

```{r}
# run cox regression based on groups
fit.surv <- survfit(Surv(survival, status) ~ Factor4_group, data = data)

ggsurv <- ggsurvplot(fit.surv,
           data = data,
           risk.table = TRUE,
           conf.int = TRUE,
           size = 1,
           pval = TRUE,
           palette = c("#0000FF", "#FB0207"),
           legend.labs = c("t_lo", "t_hi"),
           ggtheme = theme_light(),
           tables.height = 0.23,
           xlim = c(0, 5),
           pval.coord = c(0,0),
           pval.size = 4.5,
           conf.int.alpha = 0.25)

ggsurv$plot <- ggsurv$plot + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text=element_text(family="Helvetica", size=12)) +xlab("Time (years)")+ 
  theme(strip.text.x = element_blank(),
      strip.background = element_rect(colour="white", fill="white"),
      legend.position=c(.85,.85))

ggsurv$table <- ggsurv$table + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                                      text=element_text(family="Helvetica", size=12))+ggtitle("")


pdf("results/figures/survival_curves_DDHM4_scores_txn_BeatAML_median_sorafenib.pdf", width = 4, height = 5)
print(ggsurv, newpage = FALSE)
dev.off()
```



```{r}
clinics_venetoclax_azacitidine <- clinics[grepl("Venetoclax", clinics$currentRegimen),]
clinics_venetoclax_azacitidine <- clinics_venetoclax_azacitidine[grepl("Azacitidine", clinics_venetoclax_azacitidine$currentRegimen),]
nrow(clinics_venetoclax_azacitidine)
```

## Extract transcriptomics
```{r}
txn <- gex[,colnames(gex) %in% c("stable_id", clinics_venetoclax_azacitidine$patient_id)]
txn$stable_id <- rownames(txn)
condition <- data.frame(patient_id = colnames(txn))
```

## Calculate scores transcriptomics
```{r}
for (i in 1:11) {
  fname <- paste0("Factor", i)
  myFactor1 <- RNAfac[RNAfac$factor==fname,]
  pick <- txn$stable_id %in% myFactor1$feature
  x=txn[pick,drop=FALSE,]
  x=t(x[,colnames(x) != "stable_id"])
  beta <- myFactor1$value[match(colnames(x), myFactor1$feature)]
  yval=t(t(x) * beta)
  yval=rowSums(yval)
  weight <- data.frame(yval)
  colnames(weight) <- fname
  weight$patient_id <- rownames(weight)
  yval <- data.frame(patient_id = names(yval), score = yval)
  yval$group <- as.integer(yval$score > median(yval$score))+1
  colnames(yval)[2:3] <- paste0(fname, "_", colnames(yval)[2:3])
  condition <- merge(condition, yval, by = "patient_id")
}
```

```{r}
clinical =clinics[,colnames(clinics) %in% c("patient_id","status", "survival")]
data <- merge(condition, clinical, by = "patient_id")
```

```{r}
# run cox regression based on groups
fit.surv <- survfit(Surv(survival, status) ~ Factor2_group, data = data)

ggsurv <- ggsurvplot(fit.surv,
           data = data,
           risk.table = TRUE,
           conf.int = TRUE,
           size = 1,
           pval = TRUE,
           palette = c("#0000FF", "#FB0207"),
           legend.labs = c("t_lo", "t_hi"),
           ggtheme = theme_light(),
           tables.height = 0.23,
           xlim = c(0, 2),
           pval.coord = c(0,0),
           pval.size = 4.5,
           conf.int.alpha = 0.25)

ggsurv$plot <- ggsurv$plot + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), text=element_text(family="Helvetica", size=12)) +xlab("Time (years)")+ 
  theme(strip.text.x = element_blank(),
      strip.background = element_rect(colour="white", fill="white"),
      legend.position=c(.85,.85))

ggsurv$table <- ggsurv$table + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                                      text=element_text(family="Helvetica", size=12))+ggtitle("")


pdf("results/figures/survival_curves_DDHM2_scores_txn_BeatAML_median_venetoclax_azacitidine.pdf", width = 4, height = 5)
print(ggsurv, newpage = FALSE)
dev.off()
```
