---
title: "R Notebook"
output: html_notebook
---


Process :
1. Assign cells with BMMap (CellType_Broad)
NB : run BMMap with raw counts

### clean environment
```{r}
rm(list=ls())
```

### set directory
```{r setup}
    knitr::opts_knit$set(root.dir = normalizePath("/home/flomik/Code/MOFA_AML/")) 
```

```{r}
library(BoneMarrowMap)
library(Seurat)
library(dplyr)
library(ggplot2)
library(readxl)
library(data.table)
```

## Load Symphony reference
```{r}
ref <- readRDS("~/Data/CITE_seq/BoneMarrowMap-main/BoneMarrowMap_SymphonyReference.rds")
ref$save_uwot_path <- paste0("~/Data/CITE_seq/BoneMarrowMap-main/", 'BoneMarrowMap_uwot_model.uwot')
```

## load seurat object
```{r}
query <- readRDS("/home/flomik/Data/external_data_sets/BMmap/scAML_Zeng_2025_projected_cleaned.rds")
query
```

```{r}
## filter out low quality cells
# Run QC based on mapping error score, flag cells with mapping error >= 2.5 MADs above median
query <- query %>% calculate_MappingError(., reference = ref, MAD_threshold = 2.5)

# Get QC Plots
QC_plots <- plot_MappingErrorQC(query)

# Plot together - If this is too crowded, can also just call "QC_plots" aloneto display one by one
#patchwork::wrap_plots(QC_plots, ncol = 4, widths = c(0.8, 0.3, 0.8, 0.3))

# # Optional step - remove outliers with high mapping error
query <- subset(query, mapping_error_QC == 'Pass')

ncol(query)
nrow(query)
```


HSC MPP — #E8A020 (medium orange)
LMPP — #F5C842 (golden yellow)
Cycling Progenitor — #F5E070 (pale yellow)
Early GMP — #FAEFC0 (very pale yellow/cream)
Megakaryo-Erythroid lineage

MEP — #8B0000 (dark red / burgundy)
Megakaryocyte Precursor — #CC2200 (strong red)
Early Erythroid — #E8604A (salmon-red)
Late Erythroid — #F0A090 (light salmon/pink)
Innate myeloid cells

EoBasoMast Precursor — #1A3A6B (dark navy blue)
Pro-Monocyte — #2E6FAD (medium-dark blue)
Monocyte — #4A9FD4 (medium blue)
cDC — #7DC0E8 (light-medium blue)
pDC — #AADDF5 (pale blue)
Stromal — #D0EEF8 (very pale blue)
Lymphoid lineage

Early Lymphoid — #1A1A1A (near black)
Pro-B — #505050 (dark grey)
Pre-B — #787878 (medium grey)
B — #A0A0A0 (medium-light grey)
Plasma_Cell — #C8C8C8 (light grey)
NK — #E8E8E8 (very light grey)


```{r}
col <- c("#A0A0A0", # B
         "#BAE4B3", # CD4 T
         "#74C476", # CD8 T
         "#7DC0E8", # cDC
         "#F5E070", # Cycling Progenitor
         "#E8604A", # Early Erythroid
         "#FAEFC0", # Early GMP
         "#1A1A1A", # Early Lymphoid
         "#1A3A6B", # EoBasoMast Precursor
         "#E8A020", # HSC MPP
         "#F0A090", # Late Erythroid
         "#f7e391", # Late GMP
         "#F5C842", # LMPP
         "#CC2200", # Megakaryocyte Precursor
         "#8B0000", # MEP 
         "#4A9FD4", # Monocyte
         "#EDF8E9", # Naive T
         "#E8E8E8", # NK
         "#AADDF5", # pDC
         "#C8C8C8", # Plasma_Cell
         "#787878", # Pre-B
         "#505050", # Pro-B
         "#2E6FAD", # Pro-Monocyte
         "#D0EEF8", # Stromal
         "pink"
         )
```


MLL_28824 -> AEL
MLL_28855 -> AEL
```{r}
samples <- c("MLL_14666",
"MLL_16703",
"MLL_17746",
"MLL_17843",
"MLL_17844",
"MLL_28830",
"MLL_29512",
"MLL_29532",
"MLL_29538",
"MLL_30862",
"MLL_30886")
```

## remove allographs, CD4 and APL samples
```{r}
Idents(query) <- "Sample"
all_samples <- factor(query$Sample)
levels(all_samples)

query <- subset(query, idents = samples)
```

## Predict Hematopoietic Cell Types by KNN classification
```{r}
query <- predict_CellTypes(
  query_obj = query, 
  ref_obj = ref, 
  initial_label = 'initial_CellType', # celltype assignments before filtering on mapping QC
  final_label = 'predicted_CellType'  # celltype assignments with map QC failing cells assigned as NA
  )
```

# UMAP
```{r}
#DimPlot(query, reduction = 'umap_projected', group.by = c('predicted_CellType_Broad'), raster=FALSE, label=FALSE, label.size = 4, cols = col)+xlab("UMAP 1")+ylab("UMAP 2")
#ggsave(paste0("results/figures/scRNA/Zheng/score_plot_predicted_cell_types_BMmap_prospective_Zheng.pdf"), width = 9, height = 4)
```

## Pseudo time

# Predict Pseudotime values by KNN
query <- predict_Pseudotime(
  query_obj = query, 
  ref_obj = ref, 
  initial_label = 'initial_Pseudotime',  # pseudotime assignments before filtering on mapping QC
  final_label = 'predicted_Pseudotime'   # pseudotime assignments with map QC failing cells assigned as NA
)

# Visualize Hematopoietic Pseudotime in query data
FeaturePlot(query, features = c('predicted_Pseudotime'))
ggsave(paste0("results/figures/scRNA/Zheng/score_plot_predicted_pseudotime_BMmap_prospective_Zheng.pdf"), width = 5, height = 3.5)


## calculate factors for each cell
```{r}
#' Perform survival analysis for each MOFA factor
#'
#' @param file variable containing file
#' @param name name of the file
#' @param destination destination folder for file
scores_factor <- function(fname, txn, genes){
  myFactor1=RNAfac[RNAfac$factor==fname,]
  p1=c(1:nrow(myFactor1)) # for coefficients
  
  if(genes == "genes"){
    myFactor1$feature <- myFactor1$label
  }
  
  pick=txn$stable_id %in% myFactor1$feature[p1]
  #print(pick)
  mygex1=txn[pick,drop=FALSE,]
  rownames(mygex1) <- mygex1$stable_id
  mygex1=data.frame(t(mygex1[,colnames(mygex1) != "stable_id"]))
  x=as.matrix(mygex1)
  #print(dim(x))
  
  beta=myFactor1$value[match(colnames(x),myFactor1$feature)]
  beta[is.na(beta)] <- 0
  yval=t(t(x) * beta)
  yval=rowSums(yval)
  
  weight <- data.frame(yval)
  colnames(weight) <- fname
  weight$patient_id <- rownames(weight)
  
  yval <- data.frame(patient_id = names(yval), score = yval)
  yval$group <- as.integer(yval$score > median(yval$score))+1
  return(yval)
}
```

```{r}
gex <- as.data.frame(query@assays$RNA$scale.data)
gex$stable_id <- rownames(gex)
#gex$stable_id <- gsub("\\..*", "", gex$stable_id)
#gex <- aggregate(. ~ stable_id, data = gex, sum)
#rownames(gex) <- gex$stable_id
```

### import factors data
```{r}
path_factors <- "data/MOFA/AML_MOFAplus_top_factors_v04_f600 (1).xlsx"

RNA_factors=NULL
for (i in 1:11){
  factorName=paste0("Factor",i,"_RNAseq")
  cat(factorName," ")
  dat=read_excel(path_factors,sheet=factorName)
  dat=data.table(dat)
  print(nrow(dat))
  RNA_factors[[i]]=dat
}
RNAfac=do.call(rbind,RNA_factors)
RNAfac$feature=gsub(".g","",RNAfac$feature)

## set thresholds for each factor
factorNames=unique(RNAfac$factor)
topGeneThres=c(9,10,2,6,4,1,2,3,3,6,3)
names(topGeneThres)=factorNames
```

```{r}
table_scores <- data.frame(patient_id = colnames(gex))

for (factor in factorNames){
  print(factor)
  scores_F <- scores_factor(factor, gex, "genes")
  table_scores <- merge(table_scores, scores_F, by = "patient_id")
}

colnames(table_scores)[-1] <- paste0("Factor_", rep(1:11, each = 2), "_",colnames(table_scores)[-1])
colnames(table_scores) <- gsub("\\..*", "", colnames(table_scores))
colnames(table_scores)[1] <- "Cell"
```

```{r}
query@meta.data <- cbind(query@meta.data, table_scores)
```

```{r}
saveRDS(query, )
```

```{r}
query <- readRDS("/home/flomik/Data/CITE_seq/Seurat_object_predicted_class_BMmap_Zheng_with_scores_v2.rds")
```


```{r}
colnames(query@meta.data)[seq(40,60, 2)]
```

```{r}
colnames(query@meta.data)[seq(40,60, 2)] <- gsub("Factor_", "DDHM", colnames(query@meta.data)[seq(40,60, 2)])
colnames(query@meta.data)[seq(40,60, 2)] <- gsub("_score", "", colnames(query@meta.data)[seq(40,60, 2)])
colnames(query@meta.data)[seq(40,60, 2)]
```

```{r}
write.csv(query@meta.data, "results/scores_cells_and_factors_CITE_seq_Zheng_prospective.csv")
```

2,3,4,5,6,7,8
```{r}
query@meta.data$DDHM2 <- -1*query@meta.data$DDHM2
query@meta.data$DDHM3 <- -1*query@meta.data$DDHM3
query@meta.data$DDHM4 <- -1*query@meta.data$DDHM4
query@meta.data$DDHM5 <- -1*query@meta.data$DDHM5
query@meta.data$DDHM6 <- -1*query@meta.data$DDHM6
query@meta.data$DDHM7 <- -1*query@meta.data$DDHM7
query@meta.data$DDHM8 <- -1*query@meta.data$DDHM8
```

```{r}
write.csv(query@meta.data, "results/scores_cells_and_factors_CITE_seq_Zheng_prospective_corrected.csv")
```

# Visualize Hematopoietic Pseudotime in query data

```{r}
library(RColorBrewer)
```

```{r}
rev(brewer.pal(n = 11, name = "RdBu"))
```

```{r}
query_2 <- query
```

```{r}
query <- query_2
```

# Visualize Hematopoietic Pseudotime in query data

```{r}
library(RColorBrewer)
```



```{r}
query_2 <- query
```

# Plot scores
```{r}
a <- 1
for(i in seq(40,60, 2)){
  print(colnames(query@meta.data)[i])
  #query@meta.data[,i][query@meta.data[,i] > max(query@meta.data[,i])/1.7] <- max(query@meta.data[,i])/1.5
  #query@meta.data[,i][query@meta.data[,i] < -max(query@meta.data[,i])/1.7] <- -max(query@meta.data[,i])/1.5
  FeaturePlot(query, features = colnames(query@meta.data)[i], label = FALSE, repel = TRUE) +
            scale_colour_gradient2(low = "#2166AC", mid = "#fbf6f0", high = "#B2182B", midpoint = 0, space = "Lab", na.value = "grey50", guide = "colourbar")+ xlab("UMAP 1")+ylab("UMAP 2")
  ggsave(paste0("results/figures/scRNA/Zheng/review_score_plot_", colnames(query@meta.data)[i],"_scores_MOFA_Zheng.pdf"), width = 5, height = 3.5)
  #ggsave(paste0("results/figures/scRNA_2/review_score_plot_", colnames(query@meta.data)[i],"_scores_MOFA.png"), width = 4, height = 4)
  a <- a + 1
}
```

