---
title: "BoneMarrowMap KI pro scRNA"
output: html_notebook
---

Aim : calculate DDHM and compare to cell types

Process :
1. Assign cells with BoneMarrowMap (CellType_Broad)
NB : run BMMap with raw counts
2. Calculate DDHMs
3. Run correlation betweem DDHM and cell types scores
4. Check genes of interests

### clean environment
```{r}
rm(list=ls())
```

```{r}
path <- getwd()
new_path <- gsub("MOFA_AML/.*","MOFA_AML/",path)
```

### set directory (replace by your repository)
```{r setup}
    knitr::opts_knit$set(root.dir = normalizePath(new_path)) 
```


```{r}
library(BoneMarrowMap)
library(Seurat)
library(dplyr)
library(ggplot2)
library(readxl)
library(data.table)
```

## Load Symphony reference
```{r}
ref <- readRDS("data/BoneMarrowMap/BoneMarrowMap_SymphonyReference.rds")
ref$save_uwot_path <- "data/BoneMarrowMap/BoneMarrowMap_uwot_model.uwot"
```

## load seurat object
```{r}
data <- readRDS("data/input_files/CITE_seq/Wells_combined_sctransformed.rds")
data
```

## remove allographs, CD4 and APL samples
```{r}
Idents(data) <- "PatientLevel"
all_samples <- factor(data$PatientLevel)
x <- levels(all_samples)
x <- x[-c(1:3)]
data <- subset(data, idents = x)
```

## assign AML cell types to cells
```{r}
batchvar <- "PatientLevel"
# Map query dataset using Symphony (Kang et al 2021)
query <- map_Query(
    exp_query = LayerData(data, assay = "RNA", layer = "counts"), 
    metadata_query = data@meta.data,
    ref_obj = ref,
    vars = batchvar
    )

ncol(query)
```

## filter out low quality cells
```{r}
# Run QC based on mapping error score, flag cells with mapping error >= 2.5 MADs above median
query <- query %>% calculate_MappingError(., reference = ref, MAD_threshold = 2.5)

# Get QC Plots
QC_plots <- plot_MappingErrorQC(query)

# Plot together - If this is too crowded, can also just call "QC_plots" aloneto display one by one
patchwork::wrap_plots(QC_plots, ncol = 4, widths = c(0.8, 0.3, 0.8, 0.3))

# # Optional step - remove outliers with high mapping error
query <- subset(query, mapping_error_QC == 'Pass')

ncol(query)
nrow(query)
```

HSC MPP — #E8A020 (medium orange)
LMPP — #F5C842 (golden yellow)
Cycling Progenitor — #F5E070 (pale yellow)
Early GMP — #FAEFC0 (very pale yellow/cream)
Megakaryo-Erythroid lineage

MEP — #8B0000 (dark red / burgundy)
Megakaryocyte Precursor — #CC2200 (strong red)
Early Erythroid — #E8604A (salmon-red)
Late Erythroid — #F0A090 (light salmon/pink)
Innate myeloid cells

EoBasoMast Precursor — #1A3A6B (dark navy blue)
Pro-Monocyte — #2E6FAD (medium-dark blue)
Monocyte — #4A9FD4 (medium blue)
cDC — #7DC0E8 (light-medium blue)
pDC — #AADDF5 (pale blue)
Stromal — #D0EEF8 (very pale blue)
Lymphoid lineage

Early Lymphoid — #1A1A1A (near black)
Pro-B — #505050 (dark grey)
Pre-B — #787878 (medium grey)
B — #A0A0A0 (medium-light grey)
Plasma_Cell — #C8C8C8 (light grey)
NK — #E8E8E8 (very light grey)

```{r}
col <- c("#A0A0A0", # B
         "#BAE4B3", # CD4 T
         "#74C476", # CD8 T
         "#7DC0E8", # cDC
         "#F5E070", # Cycling Progenitor
         "#E8604A", # Early Erythroid
         "#FAEFC0", # Early GMP
         "#1A1A1A", # Early Lymphoid
         "#1A3A6B", # EoBasoMast Precursor
         "#E8A020", # HSC MPP
         "#F0A090", # Late Erythroid
         "#f7e391", # Late GMP
         "#F5C842", # LMPP
         "#CC2200", # Megakaryocyte Precursor
         "#8B0000", # MEP 
         "#4A9FD4", # Monocyte
         "#EDF8E9", # Naive T
         "#E8E8E8", # NK
         "#AADDF5", # pDC
         "#C8C8C8", # Plasma_Cell
         "#787878", # Pre-B
         "#505050", # Pro-B
         "#2E6FAD", # Pro-Monocyte
         "#D0EEF8" # Stromal
         )
```



## Predict Hematopoietic Cell Types by KNN classification
```{r}
query <- predict_CellTypes(
  query_obj = query, 
  ref_obj = ref, 
  initial_label = 'initial_CellType', # celltype assignments before filtering on mapping QC
  final_label = 'predicted_CellType'  # celltype assignments with map QC failing cells assigned as NA
)
```


```{r}
DimPlot(query, reduction = 'umap_projected', group.by = c('predicted_CellType_Broad'), 
        raster=FALSE, label=FALSE, label.size = 4, cols = col) + xlab("UMAP 1")+ylab("UMAP 2")

ggsave(paste0("results/figures/scRNA/KI/score_plot_predicted_cell_types_BMmap_prospective_KI.pdf"), width = 9, height = 4)
```

# Predict Pseudotime values by KNN
query <- predict_Pseudotime(
  query_obj = query, 
  ref_obj = ref, 
  initial_label = 'initial_Pseudotime',  # pseudotime assignments before filtering on mapping QC
  final_label = 'predicted_Pseudotime'   # pseudotime assignments with map QC failing cells assigned as NA
)

# Visualize Hematopoietic Pseudotime in query data
FeaturePlot(query, features = c('predicted_Pseudotime'))

ggsave(paste0("results/figures/scRNA/KI/score_plot_predicted_pseudotime_BMmap_prospective_KI.pdf"), width = 5, height = 3.5)



```{r}
#saveRDS(query, "/home/flomik/Data/CITE_seq/Seurat_object_predicted_class_BMmap_KI_v2.rds")
```

## calculate factors for each cell
```{r}
#' Perform survival analysis for each MOFA factor
#'
#' @param file variable containing file
#' @param name name of the file
#' @param destination destination folder for file
scores_factor <- function(fname, txn, genes){
  myFactor1=RNAfac[RNAfac$factor==fname,]
  p1=c(1:nrow(myFactor1)) # for coefficients
  
  if(genes == "genes"){
    myFactor1$feature <- myFactor1$label
  }
  
  pick=txn$stable_id %in% myFactor1$feature[p1]
  #print(pick)
  mygex1=txn[pick,drop=FALSE,]
  rownames(mygex1) <- mygex1$stable_id
  mygex1=data.frame(t(mygex1[,colnames(mygex1) != "stable_id"]))
  x=as.matrix(mygex1)
  #print(dim(x))
  
  beta=myFactor1$value[match(colnames(x),myFactor1$feature)]
  beta[is.na(beta)] <- 0
  yval=t(t(x) * beta)
  yval=rowSums(yval)
  
  weight <- data.frame(yval)
  colnames(weight) <- fname
  weight$patient_id <- rownames(weight)
  
  yval <- data.frame(patient_id = names(yval), score = yval)
  yval$group <- as.integer(yval$score > median(yval$score))+1
  return(yval)
}
```

```{r}
gex <- as.data.frame(query@assays$RNA$scale.data)
gex$stable_id <- rownames(gex)
#gex$stable_id <- gsub("\\..*", "", gex$stable_id)
#gex <- aggregate(. ~ stable_id, data = gex, sum)
#rownames(gex) <- gex$stable_id
```

### import factors data
```{r}
path_factors <- "data/MOFA/AML_MOFAplus_top_factors_v04_f600 (1).xlsx"

RNA_factors=NULL
for (i in 1:11){
  factorName=paste0("Factor",i,"_RNAseq")
  cat(factorName," ")
  dat=read_excel(path_factors,sheet=factorName)
  dat=data.table(dat)
  print(nrow(dat))
  RNA_factors[[i]]=dat
}

RNAfac=do.call(rbind,RNA_factors)
RNAfac$feature=gsub(".g","",RNAfac$feature)

## set thresholds for each factor
factorNames=unique(RNAfac$factor)
topGeneThres=c(9,10,2,6,4,1,2,3,3,6,3)
names(topGeneThres)=factorNames
```

```{r}
table_scores <- data.frame(patient_id = colnames(gex))

for (factor in factorNames){
  print(factor)
  scores_F <- scores_factor(factor, gex, "genes")
  table_scores <- merge(table_scores, scores_F, by = "patient_id")
}

colnames(table_scores)[-1] <- paste0("Factor_", rep(1:11, each = 2), "_",colnames(table_scores)[-1])
colnames(table_scores) <- gsub("\\..*", "", colnames(table_scores))
colnames(table_scores)[1] <- "Cell"
```

```{r}
query@meta.data <- cbind(query@meta.data, table_scores)
```

```{r}
#saveRDS(query, "/home/flomik/Data/CITE_seq/Seurat_object_predicted_class_BMmap_KI_with_scores_v2.rds")
```

```{r}
#query <- readRDS("/home/flomik/Data/CITE_seq/Seurat_object_predicted_class_BMmap_KI_with_scores_v2.rds")
```

```{r}
DimPlot(query, reduction = 'umap_projected', group.by = c('PatientLevel'), 
        raster=FALSE, label=FALSE, label.size = 4) + xlab("UMAP 1")+ylab("UMAP 2")

ggsave(paste0("results/figures/scRNA/KI/UMAP_patients_KI_pro_after_BMMAP.pdf"), width = 6, height = 4)
```

```{r}
colnames(query@meta.data)[seq(60,81, 2)]
colnames(query@meta.data)[seq(60,81, 2)] <- gsub("Factor_", "DDHM", colnames(query@meta.data)[seq(60,81, 2)])
colnames(query@meta.data)[seq(60,81, 2)] <- gsub("_score", "", colnames(query@meta.data)[seq(60,81, 2)])
colnames(query@meta.data)[seq(60,81, 2)]
```

```{r}
write.csv(table_scores, "results/scRNA/scores_cells_and_factors_CITE_seq_KI_prospective.csv")
```


## change 4, 5 and 8
```{r}
query@meta.data$DDHM4 <- -1*query@meta.data$DDHM4
query@meta.data$DDHM5 <- -1*query@meta.data$DDHM5
query@meta.data$DDHM8 <- -1*query@meta.data$DDHM8
```


# Visualize Hematopoietic Pseudotime in query data

```{r}
library(RColorBrewer)
```



```{r}
query_2 <- query
```

# Plot scores
```{r}
a <- 1
for(i in seq(60,81, 2)){
  print(colnames(query@meta.data)[i])
  FeaturePlot(query, features = colnames(query@meta.data)[i], label = FALSE, repel = TRUE) +
            scale_colour_gradient2(low = "#2166AC", mid = "#fbf6f0", high = "#B2182B", midpoint = 0, space = "Lab", na.value = "grey50", guide = "colourbar")+ xlab("UMAP 1")+ylab("UMAP 2")
  ggsave(paste0("results/figures/scRNA/KI/review_score_plot_", colnames(query@meta.data)[i],"_scores_MOFA.pdf"), width = 5, height = 3.5)
  a <- a + 1
}
```

Heatmap surface markers + factors values by cell

```{r}
factors <- query@meta.data[,c(which(colnames(query@meta.data) %in% "predicted_CellType_Broad"), seq(60,81, 2))]
factors$Cell <- rownames(factors)
surface_markers <- data.frame(t(data@assays$ADT$scale.data))
surface_markers$Cell <- rownames(surface_markers)
data <- merge(factors, surface_markers, by = "Cell")
```

```{r}
library(ComplexHeatmap)
```

```{r}
data_heatmap <- data
data_heatmap <- data_heatmap[!is.na(data_heatmap$predicted_CellType_Broad),]
data_type <- colnames(data)[-c(1, 2)]
data_type <- gsub("DDHM.*", "DDHMs", data_type)
data_type[!grepl("DDHM", data_type)] <- "Surface_proteins"
data_heatmap[,3:13] <- scale(data_heatmap[,3:13])
data_heatmap[,14:ncol(data_heatmap)] <- scale(data_heatmap[,14:ncol(data_heatmap)])
colnames(data_heatmap) <- gsub(".TotalA", "", colnames(data_heatmap))

x <- names(table(data_heatmap$predicted_CellType_Broad)[table(data_heatmap$predicted_CellType_Broad) < 100])
data_heatmap$predicted_CellType_Broad[data_heatmap$predicted_CellType_Broad %in% x] <- "Others"

pdf("results/figures/scRNA/KI/Heatmap_KI_CITE_seq_cell_types_BMmap_surface_proteins_factors.pdf", width = 11)
Heatmap(t(data_heatmap[,-c(1,2)]), name = "Z-score", column_split = data_heatmap$predicted_CellType_Broad, row_split =  data_type, show_row_names = TRUE, show_column_names = FALSE, row_title_rot = 0, column_title_rot = 60)
dev.off()
```

## check genes of interest
```{r}
list_genes <- read.delim("/home/flomik/Code/MOFA_AML/data/Genes/genes_UMAP_CITE_seq_KI.txt")
list_genes <- list_genes$Genes
```

TBXAX1, GATA4, PRUNE1 -_> not found
```{r}
FeaturePlot(query, features = list_genes, label = FALSE, repel = TRUE, cols = c("#fbf6f0","#B2182B"))+ xlab("UMAP 1")+ylab("UMAP 2")
ggsave(paste0("results/figures/scRNA/KI/UMAP_genes_list_genes_Tom_KI_pro.pdf"), width = 12, height = 10)
```

GATA2 x GDI1
GATA2 x MPP1
GATA2 x PRUNE1 --> not found
GATA2 x EMILIN2 (edited)
```{r}
pair_1 <- list(c("GATA2", "GDI1"))
query <- AddModuleScore(object = query, features = pair_1, name = "GATA2_GDI1_score")

pair_2 <- list(c("GATA2", "MPP1"))
query <- AddModuleScore(object = query, features = pair_2, name = "GATA2_MPP1_score")


pair_3 <- list(c("GATA2", "EMILIN2"))
query <- AddModuleScore(object = query, features = pair_3, name = "GATA2_EMILIN2_score")
```


```{r}
FeaturePlot(object = query, features = c("GATA2_GDI1_score1", "GATA2_MPP1_score1", "GATA2_EMILIN2_score1"), label = FALSE, repel = TRUE, cols = c("#fbf6f0","#B2182B"))+ xlab("UMAP 1")+ylab("UMAP 2")
ggsave(paste0("results/figures/scRNA/KI/UMAP_genes_list_genes_Tom_KI_pro_pairs_of_genes.pdf"), width = 6, height = 5)            
```

