---
title: "DDHM"
output: html_document
date: "2025-10-10"
---

Author : Nona Struyf

```{r setup, include=FALSE,  results = 'hide'}
library(biomaRt)
library(Seurat)
library(dplyr)
library(ggplot2)
library(ggrepel)
library(tidyverse)
library(symphony)
library(ggpubr)
library(patchwork)
library(BoneMarrowMap)
library(pheatmap)
library(RColorBrewer)
library(ComplexHeatmap)
library(grid)
library(AUCell)
library(data.table)

```


## scRNAseq data analysis

Using the BoneMarrowMap pipeline, we can assign reference cell types to each cell in the dataset. First all data is loaded and formatted before mapping the data to the BoneMarrowMap.


```{r,  results = 'hide', message=FALSE, warning=FALSE}

ref <- readRDS("~/Projects/LSC_immune/data/external/BMM/BoneMarrowMap_SymphonyReference.rds")
ref$save_uwot_path <- "~/Projects/LSC_immune/data/external/BMM/BoneMarrowMap_uwot_model.uwot"

OK_data <- readRDS("~/Projects/LSC_immune/data/raw/3a_data_cite_seq.rds")
metadata_OK <- read.csv("~/Projects/LSC_immune/data/raw/3b_data_cite_seq_meta_data.csv", header = TRUE)
gene_mat <- as.matrix(OK_data)

mart <- useEnsembl(biomart = "genes", dataset = "hsapiens_gene_ensembl", version = 108)
gene_map <- getBM(
  attributes = c("ensembl_gene_id", "hgnc_symbol"),
  filters = "ensembl_gene_id",
  values = rownames(gene_mat),
  mart = mart
)
gene_map <- gene_map[gene_map$hgnc_symbol != "", ]
gene_map <- gene_map[!duplicated(gene_map$ensembl_gene_id), ]
new_names <- gene_map$hgnc_symbol[match(rownames(gene_mat), gene_map$ensembl_gene_id)]
rownames(gene_mat) <- ifelse(is.na(new_names), rownames(gene_mat), new_names)
dup_genes <- duplicated(rownames(gene_mat))
gene_mat <- gene_mat[!dup_genes, ]

seurat_OK <- CreateSeuratObject(counts = gene_mat, project = "sc_OK")
rownames(metadata_OK) <- metadata_OK$X
seurat_OK <- AddMetaData(seurat_OK, metadata = metadata_OK)
seurat_OK <- JoinLayers(seurat_OK)

query_ok <- seurat_OK

rownames(query_ok[['RNA']]@layers[['counts']]) <- rownames(query_ok[['RNA']])
colnames(query_ok[['RNA']]@layers[['counts']]) <- colnames(query_ok)
batchvar <- 'Sample'

DefaultAssay(query_ok) <- "RNA"


# run mapping
query_ok <- map_Query(
  exp_query = query_ok[['RNA']]@layers[['counts']],
  metadata_query = query_ok@meta.data,
  ref_obj = ref,
  vars = batchvar
)

# QC steps
query_ok <- query_ok %>% calculate_MappingError(., reference = ref, MAD_threshold = 2.5) 

```

# Prediction of cell types based on BoneMarrowMap

Cell types are assigned to cells that passed the QC

```{r,  results = 'hide'}
query_ok <- predict_CellTypes(
  query_obj = query_ok,
  ref_obj = ref,
  initial_label = 'initial_CellType', # celltype assignment before filtering on mapping QC
  final_label = 'predicted_CellType'  # celltype assignments with map QC failing cells assigned as NA
)

```

Clustering

```{r,  results = 'hide', message=FALSE, warning=FALSE, echo=FALSE}
query_ok <- FindVariableFeatures(query_ok, selection.method = "vst", nfeatures = 2000)
query_ok <- RunPCA(query_ok, features = VariableFeatures(query_ok))
query_ok <- RunUMAP(query_ok, dims = 1:20) 
query_ok <- FindNeighbors(query_ok, dims = 1:20)
query_ok <- FindClusters(query_ok, resolution = 0.5)
```

Cell type UMAP

```{r}

custom_colors_type <- c(
  "HSC MPP"               = "#1B9E77",
  "LMPP"                  = "#33A65C",
  "Cycling Progenitor"    = "#66B34E",
  "Early Lymphoid"        = "#8E63A2",
  "Pro-B"                 = "#A770B5",
  "Pre-B"                 = "#C77CFF",
  "B"                     = "#E07BEF",
  "Plasma Cell"           = "#FF6BCF",
  "Naive T"               = "#E41A1C",
  "CD4 Memory T"          = "#FF5733",
  "CD8 Memory T"          = "#FF7043",
  "NK"                    = "#FF8A65",
  "MEP"                   = "#FFD92F",
  "Early Erythroid"       = "#FFC000",
  "Late Erythroid"        = "#E6AC00",
  "Megakaryocyte Precursor" = "#CC9900",
  "Early GMP"             = "#00A896",
  "Late GMP"              = "#00C2C7",
  "Pro-Monocyte"          = "#00CED1",
  "Monocyte"              = "#40E0D0",
  "pDC"                   = "#377EB8",
  "cDC"                   = "#4F94CD",
  "EoBasoMast Precursor"  = "#A39BD8",
  "Stromal"               = "#7CAE00"
)

my_levels <- c(
  "HSC MPP", "LMPP", "Cycling Progenitor",
  "Early Lymphoid", "Pro-B", "Pre-B", "B", "Plasma Cell",
  "Naive T", "CD4 Memory T", "CD8 Memory T", "NK",
  "MEP", "Early Erythroid", "Late Erythroid", "Megakaryocyte Precursor",
  "Early GMP", "Late GMP", "Pro-Monocyte", "Monocyte",
  "pDC", "cDC",
  "EoBasoMast Precursor",
  "Stromal"
)

query_ok$predicted_CellType_Broad <- factor(query_ok$predicted_CellType_Broad, levels = my_levels)

DimPlot(subset(query_ok, mapping_error_QC == 'Pass'), reduction = 'umap_projected', group.by = c('predicted_CellType_Broad'), cols = custom_colors_type,
        raster = FALSE, label = FALSE, label.size = 4) + labs(x = "UMAP1", y = "UMAP2") + ggtitle("Predicted cell type") + theme(
    plot.title = element_text(size = 12, face = "bold"),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 12),
    legend.text = element_text(size = 12),
    strip.text = element_text(size = 12)
  )
```


# Custom gene set enrichment

Calculate enrichment of the different gene sets for each single cell

```{r}
ery_mk_genes <- c("HBA1", "HBA2", "HBG1", "HBG2", "ALAS2", "DMTN", "TRIM58", "FAXDC2", "SELENBP1", 
  "TAL1", "SLC14A1", "OSBP2", "PPBP", "ITGB3", "GP1BA", "PARVB", "RAB27B", 
  "THBS1", "PDE3A", "TNIK", "SKAP1", "KRT1", "ARHGAP6", "GATA1", "TUBB2B")

gene_set1 <- c("TC2N", "ARHGAP6", "TNIK", "TAL1", "SLC14A1", "XK")
gene_set2 <- c("PPBP", "GP1BA", "PARVB", "C2orf88", "THBS1", "KALRN", "ITGB3", "RAB27B")
gene_set_combined <- c("TC2N", "ARHGAP6", "TNIK", "TAL1", "SLC14A1", "XK", "PPBP", "GP1BA", "PARVB", "C2orf88", "THBS1", "KALRN", "ITGB3", "RAB27B")
gene_sets <- list(Set1 = gene_set1, Set2 = gene_set2, Set3 = gene_set_combined)


exprMatrix <- GetAssayData(query_ok, assay = "RNA", slot = "data")

cells_rankings <- AUCell_buildRankings(exprMatrix, plotStats = FALSE, verbose = FALSE)
cells_AUC <- AUCell_calcAUC(gene_sets, cells_rankings)

query_ok$Set1_score <- as.numeric(getAUC(cells_AUC)["Set1", ])
query_ok$Set2_score <- as.numeric(getAUC(cells_AUC)["Set2", ])
query_ok$Set3_score <- as.numeric(getAUC(cells_AUC)["Set3", ])
query_ok <- subset(query_ok, subset = !is.na(predicted_CellType_Broad))


df_scores <- query_ok@meta.data %>%
  dplyr::select(predicted_CellType_Broad, Set1_score, Set2_score, Set3_score) %>%
  dplyr::group_by(predicted_CellType_Broad) %>%
  dplyr::summarise(
    across(starts_with("Set"), ~ mean(.x, na.rm = TRUE))
  ) %>%
  dplyr::arrange(desc(Set1_score))

```

Visualize enrichment of the three sets in waterfall plots

```{r}

set1 <- ggplot(df_scores, aes(x = reorder(predicted_CellType_Broad, Set1_score), y = Set1_score)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(x = "BoneMarrowMap Cell Type (Broad)", y = "Mean Enrichment Score", title = "Set1 enrichment") +
  theme_bw() +
  theme(
    plot.title = element_text(size = 12, face = "bold"),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 12),
    legend.text = element_text(size = 12),
    strip.text = element_text(size = 12)
  )

set2 <- ggplot(df_scores, aes(x = reorder(predicted_CellType_Broad, Set2_score), y = Set2_score)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(x = "", y = "Mean Enrichment Score", title = "Set2 enrichment") +
  theme_bw() +
  theme(
    plot.title = element_text(size = 12, face = "bold"),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 12),
    legend.text = element_text(size = 12),
    strip.text = element_text(size = 12)
  )

set3 <- ggplot(df_scores, aes(x = reorder(predicted_CellType_Broad, Set3_score), y = Set3_score)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(x = "", y = "Mean Enrichment Score", title = "Combined sets enrichment") +
  theme_bw() +
  theme(
    plot.title = element_text(size = 12, face = "bold"),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 12),
    legend.text = element_text(size = 12),
    strip.text = element_text(size = 12)
  )


set1 + set2 + set3 + plot_layout(ncol = 3, widths = c(0.8, 0.8))

```
 Or project enrichment on the UMAP
 
```{r}
set1 <- FeaturePlot(query_ok, features = "Set1_score", reduction = "umap_projected") +
  labs(x = "UMAP1", y = "UMAP2") +
  scale_color_viridis_c() +
  ggtitle("Set1 enrichment")

set2 <- FeaturePlot(query_ok, features = "Set2_score", reduction = "umap_projected") +
  scale_color_viridis_c() +
  labs(x = "UMAP1", y = "UMAP2") +
  ggtitle("Set2 enrichment")


set3 <- FeaturePlot(query_ok, features = "Set3_score", reduction = "umap_projected") +
  scale_color_viridis_c() +
  labs(x = "UMAP1", y = "UMAP2") +
  ggtitle("Combined enrichment")

set1 + set2 + plot_layout(ncol = 2, widths = c(0.8, 0.8))

set3 + plot_layout(ncol = 1, widths = c(0.8, 0.8))

```
permutation test of signature in Ery/MEK assigned cells

```{r}
gene_set1 <- c("TC2N", "ARHGAP6", "TNIK", "TAL1", "SLC14A1", "XK")
gene_set2 <- c("PPBP", "GP1BA", "PARVB", "C2orf88", "THBS1", "KALRN", "ITGB3", "RAB27B")
gene_set_combined <- c(gene_set1, gene_set2)

query_ok <- subset(query_ok, mapping_error_QC == 'Pass')

exprMatrix <- GetAssayData(query_ok, assay = "RNA", slot = "data")

target_celltypes <- c("MEP", "Early Erythroid", "Late Erythroid", "Megakaryocyte Precursor")

all_genes <- rownames(exprMatrix)

calculate_enrichment_metric <- function(query_ok, score_column, target_types) {
  df <- query_ok@meta.data %>%
    filter(!is.na(predicted_CellType_Broad)) %>%
    mutate(is_target = predicted_CellType_Broad %in% target_types) %>%
    group_by(is_target) %>%
    summarise(mean_score = mean(.data[[score_column]], na.rm = TRUE))
  
  target_mean <- df$mean_score[df$is_target == TRUE]
  nontarget_mean <- df$mean_score[df$is_target == FALSE]
  
  return(target_mean - nontarget_mean)
}

observed_diff_set1 <- calculate_enrichment_metric(query_ok, "Set1_score", target_celltypes)
observed_diff_set2 <- calculate_enrichment_metric(query_ok, "Set2_score", target_celltypes)
observed_diff_combined <- calculate_enrichment_metric(query_ok, "Set3_score", target_celltypes)

set.seed(123)
n_permutations <- 1000

null_diffs_set1 <- numeric(n_permutations)
null_diffs_set2 <- numeric(n_permutations)
null_diffs_combined <- numeric(n_permutations)


for (i in 1:n_permutations) {
  random_set1 <- sample(all_genes, size = length(gene_set1))
  random_set2 <- sample(all_genes, size = length(gene_set2))
  random_combined <- c(random_set1, random_set2)
  
  random_gene_sets <- list(
    Random1 = random_set1,
    Random2 = random_set2,
    Random_combined = random_combined
  )
  
  random_AUC <- AUCell_calcAUC(random_gene_sets, cells_rankings, verbose = FALSE)
  
  auc_matrix <- getAUC(random_AUC)
  cell_names <- colnames(query_ok)
  
  temp_df <- query_ok@meta.data %>%
    mutate(
      cell_id = rownames(.),
      random1 = auc_matrix["Random1", cell_id],
      random2 = auc_matrix["Random2", cell_id],
      random_combined = auc_matrix["Random_combined", cell_id]
    )
  
  null_diffs_set1[i] <- temp_df %>%
    filter(!is.na(predicted_CellType_Broad)) %>%
    mutate(is_target = predicted_CellType_Broad %in% target_celltypes) %>%
    group_by(is_target) %>%
    summarise(mean_score = mean(random1, na.rm = TRUE)) %>%
    summarise(diff = mean_score[is_target == TRUE] - mean_score[is_target == FALSE]) %>%
    pull(diff)
  
  null_diffs_set2[i] <- temp_df %>%
    filter(!is.na(predicted_CellType_Broad)) %>%
    mutate(is_target = predicted_CellType_Broad %in% target_celltypes) %>%
    group_by(is_target) %>%
    summarise(mean_score = mean(random2, na.rm = TRUE)) %>%
    summarise(diff = mean_score[is_target == TRUE] - mean_score[is_target == FALSE]) %>%
    pull(diff)
  
  null_diffs_combined[i] <- temp_df %>%
    filter(!is.na(predicted_CellType_Broad)) %>%
    mutate(is_target = predicted_CellType_Broad %in% target_celltypes) %>%
    group_by(is_target) %>%
    summarise(mean_score = mean(random_combined, na.rm = TRUE)) %>%
    summarise(diff = mean_score[is_target == TRUE] - mean_score[is_target == FALSE]) %>%
    pull(diff)
  
  if (i %% 100 == 0) cat("Completed", i, "permutations\n")
}

p_value_set1 <- sum(null_diffs_set1 >= observed_diff_set1) / n_permutations
p_value_set2 <- sum(null_diffs_set2 >= observed_diff_set2) / n_permutations
p_value_combined <- sum(null_diffs_combined >= observed_diff_combined) / n_permutations

cat("\n=== PERMUTATION TEST RESULTS ===\n")
cat("\nSet1 (Erythroid):\n")
cat("  Observed enrichment difference:", round(observed_diff_set1, 4), "\n")
cat("  Mean random difference:", round(mean(null_diffs_set1), 4), "\n")
cat("  P-value:", p_value_set1, "\n")
cat("  Z-score:", round((observed_diff_set1 - mean(null_diffs_set1)) / sd(null_diffs_set1), 2), "\n")

cat("\nSet2 (Megakaryocyte):\n")
cat("  Observed enrichment difference:", round(observed_diff_set2, 4), "\n")
cat("  Mean random difference:", round(mean(null_diffs_set2), 4), "\n")
cat("  P-value:", p_value_set2, "\n")
cat("  Z-score:", round((observed_diff_set2 - mean(null_diffs_set2)) / sd(null_diffs_set2), 2), "\n")

cat("\nCombined Set:\n")
cat("  Observed enrichment difference:", round(observed_diff_combined, 4), "\n")
cat("  Mean random difference:", round(mean(null_diffs_combined), 4), "\n")
cat("  P-value:", p_value_combined, "\n")
cat("  Z-score:", round((observed_diff_combined - mean(null_diffs_combined)) / sd(null_diffs_combined), 2), "\n")

plot_data <- data.frame(
  Difference = c(null_diffs_set1, null_diffs_set2, null_diffs_combined),
  GeneSet = rep(c("Set1 (Erythroid)", "Set2 (Megakaryocyte)", "Combined"), 
                each = n_permutations)
)

observed_data <- data.frame(
  Observed = c(observed_diff_set1, observed_diff_set2, observed_diff_combined),
  GeneSet = c("Set1 (Erythroid)", "Set2 (Megakaryocyte)", "Combined"),
  PValue = c(p_value_set1, p_value_set2, p_value_combined)
)

p <- ggplot(plot_data, aes(x = Difference)) +
  geom_histogram(bins = 50, fill = "grey70", alpha = 0.7) +
  geom_vline(data = observed_data, aes(xintercept = Observed), 
             color = "red", linewidth = 1, linetype = "dashed") +
  geom_text(data = observed_data, 
            aes(x = Observed, y = Inf, 
                label = paste0("Observed = ", round(Observed, 3), 
                              "\np = ", ifelse(PValue == 0, paste0("< ", 1/n_permutations), PValue))),
            vjust = 1.2, hjust = -0.05, color = "red", size = 3.5) +
  facet_wrap(~ GeneSet, scales = "free", ncol = 1) +
  labs(title = "Permutation Test: Ery/MK Gene Set Enrichment",
       subtitle = paste("Target cell types:", paste(target_celltypes, collapse = ", "), 
                       "\n", n_permutations, "random gene sets"),
       x = "Enrichment Difference (Target - Non-target cell types)",
       y = "Count") +
  theme_bw() +
  theme(
    plot.title = element_text(size = 12, face = "bold"),
    axis.title = element_text(size = 11),
    axis.text = element_text(size = 10),
    strip.text = element_text(size = 11, face = "bold")
  )

results_df <- data.frame(
  GeneSet = c("Set1", "Set2", "Combined"),
  N_genes = c(length(gene_set1), length(gene_set2), length(gene_set_combined)),
  Observed_difference = c(observed_diff_set1, observed_diff_set2, observed_diff_combined),
  Mean_random_difference = c(mean(null_diffs_set1), mean(null_diffs_set2), mean(null_diffs_combined)),
  SD_random = c(sd(null_diffs_set1), sd(null_diffs_set2), sd(null_diffs_combined)),
  Z_score = c(
    (observed_diff_set1 - mean(null_diffs_set1)) / sd(null_diffs_set1),
    (observed_diff_set2 - mean(null_diffs_set2)) / sd(null_diffs_set2),
    (observed_diff_combined - mean(null_diffs_combined)) / sd(null_diffs_combined)
  ),
  P_value = c(p_value_set1, p_value_set2, p_value_combined)
)

print(results_df)
```

## Bulk RNAseq data

Wilcox test comparing gene signature between DDHM5 low and high patients

```{r}
expr_data <- fread("~/Projects/LSC_immune/data/raw/ComBat_integrated_intron_exon_counts_gene_names.txt.gz")
expr_mat <- as.data.frame(expr_data)
rownames(expr_mat) <- expr_mat$V1
expr_mat[[1]] <- NULL
expr_mat <- as.matrix(expr_mat)
expr_mat <- log2(expr_mat + 1)


df_prot = read.table("~/Projects/DDHM/data/raw/GSEA_proteomics_dataset_R.txt", stringsAsFactors = FALSE,
                     header = TRUE, quote = "", comment.char = "",sep = "\t")

if (any(duplicated(df_prot$Name))) {
  df_prot <- df_prot[!duplicated(df_prot$Name), ]
}
df_filtered <- df_prot[rowMeans(!is.na(df_prot)) > 0.5, ]
rownames(df_filtered) <- df_filtered$Name
df_filtered$Name <- NULL

ddhm_vals <- read.table("~/Projects/DDHM/data/processed/DDHMvals.txt", stringsAsFactors = FALSE, header = TRUE, quote = "", comment.char = "", sep = "\t")
rownames(ddhm_vals) <- ddhm_vals$sample
ddhm_vals$sample <- NULL
ddhm_vals <- t(ddhm_vals)

ery_mk_genes <- c("HBA1", "HBA2", "HBG1", "HBG2", "ALAS2", "DMTN", "TRIM58", "FAXDC2", "SELENBP1", 
  "TAL1", "SLC14A1", "OSBP2", "PPBP", "ITGB3", "GP1BA", "PARVB", "RAB27B", 
  "THBS1", "PDE3A", "TNIK", "SKAP1", "KRT1", "ARHGAP6", "GATA1", "TUBB2B")


ery_mek_genes_mat <- intersect(ery_mk_genes, rownames(expr_mat))
sig_scores <- colMeans(expr_mat[ery_mk_genes, ], na.rm = TRUE)

ddhm5 <- as.numeric(ddhm_vals["DDHM5", ])
names(ddhm5) <- colnames(ddhm_vals)
common_patients <- intersect(names(sig_scores), names(ddhm5))
sig_scores2 <- sig_scores[common_patients]
ddhm5_2 <- ddhm5[common_patients]

plot_df <- data.frame(
  patient = common_patients,
  signature = as.numeric(sig_scores2),
  ddhm5 = as.numeric(ddhm5_2)
)

plot_df$group <- ifelse(
  plot_df$ddhm5 < median(plot_df$ddhm5, na.rm = TRUE),
  "DDHM5_low",
  "DDHM5_high"
)


table(plot_df$group)
wilcox.test(signature ~ group, data = plot_df)

ggplot(plot_df, aes(group, signature)) +
  geom_boxplot() +
  geom_jitter(width = 0.1, size = 1) +
  stat_compare_means(
    comparisons = list(c("DDHM5_low", "DDHM5_high")),
    method = "wilcox.test",
    label = "p.format",
    bracket.size = 0.5,
    tip.length = 0.01,
    size = 4
  ) +
  theme_classic() +
  labs(
    title = "MMS signature in DDHM5",
    y = "Signature score",
    x = ""
  )

```

Permutation test of the gene signature in DDHM5 low/high patients

```{r}

set.seed(123)

all_genes <- rownames(expr_mat)
ngenes <- length(ery_mk_genes)

nperm <- 10000
null_stats <- numeric(nperm)

for(i in 1:nperm) {

  random_genes <- sample(all_genes, ngenes)

  random_score <- colMeans(
    expr_mat[random_genes, , drop = FALSE],
    na.rm = TRUE
  )

  random_score <- random_score[plot_df$patient]

  low_vals <- random_score[plot_df$group == "DDHM5_low"]
  high_vals <- random_score[plot_df$group == "DDHM5_high"]

  null_stats[i] <- mean(low_vals, na.rm = TRUE) -
                   mean(high_vals, na.rm = TRUE)
}

null_stats <- null_stats[!is.na(null_stats)]

observed_stat <- mean(
  plot_df$signature[plot_df$group == "DDHM5_low"],
  na.rm = TRUE
) -
mean(
  plot_df$signature[plot_df$group == "DDHM5_high"],
  na.rm = TRUE
)

empirical_p <- mean(abs(null_stats) >= abs(observed_stat))

```

```{r}
perm_df <- data.frame(null_stats = null_stats)
ggplot(perm_df, aes(null_stats)) +
  geom_histogram(bins = 50) +
  geom_vline(
    xintercept = observed_stat,
    color = "red",
    linewidth = 1
  ) +
  annotate(
    "text",
    x = observed_stat,
    y = Inf,
    label = paste0("Observed\nEmpirical p = <0.0001"),
    vjust = 2,
    hjust = -0.1,
    color = "red"
  ) +
  theme_classic() +
  labs(
    title = "Permutation test",
    x = "Difference in signature score",
    y = "Count"
  ) +
  xlim(-0.5,2.5)

```

Spearman correlation with gene signature and all DDHMs

```{r}
all_corrs <- apply(ddhm_vals, 1, function(x) {
  cor(x[plot_df$patient],
      plot_df$signature,
      method = "spearman",
      use = "complete.obs")
})

all_corrs

all_pvals <- apply(ddhm_vals, 1, function(x) {
  cor.test(
    x[plot_df$patient],
    plot_df$signature,
    method = "spearman"
  )$p.value
})

all_pvals

```

```{r}

corr_df <- data.frame(
  DDHM = names(all_corrs),
  rho = as.numeric(all_corrs)
)
corr_df$highlight <- ifelse(corr_df$DDHM == "DDHM5", "DDHM5", "Other")
ggplot(corr_df, aes(x = reorder(DDHM, rho), y = rho, fill = highlight)) +
  geom_bar(stat = "identity") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_fill_manual(values = c("DDHM5" = "red", "Other" = "grey70")) +
  theme_classic() +
  labs(
    x = "",
    y = "Spearman correlation (MMS signature)"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )

```

